---
title: "YNP_seasons_NDVI_snowpack"
author: "Molly Caldwell"
date: "3/3/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#packages
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(parallel)
library(lubridate)
```

#Load data
GPS data: cleaned GPS data of 5 YNP ungulate species combined
  -bison, bighorn, and mule deer have 2 hour fixes
  -elk (1-hr fixes) and pronghorn (4-hr fixes) sampled to 2 hour fixes
  
Defined polygon 50 km around all GPS data for cropping the raster environmental data.

```{r}
#load all spp GPS collar data, all years/seasons
gps_data <- readRDS("~/UWyo/PhD project/HOTR-proj/Data/GPS data/Cleaned/2021/allspp_elk_ph_2hr_cleaned_notseasonal_11.2021.rds")

#define study area polygon that's 50km around edges around data 
studyareapoly <- gps_data %>% 
  extent() %>% 
  extend(c(50000, 50000, 50000, 50000)) %>%  # add 50km buffer
  extent() %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_sf(data.frame(id="studyarea")) #add a dataframe, so you can write it to file 

st_crs(studyareapoly) <- st_crs(gps_data)   # set the projection
```

#NDVI/brown down peak dates

Use the NDVI SE fit from Jerod's GIS folder to create rasters that represent
spring start of greenup, the max NDVI date, the mean peak NDVI date, and
the mean peak brown down date across the study area.

```{r}
# bring in SE of the fit
SE <- stack("Z:/MODIS_NDVI/Bischof_calculations/dlcparams.grd")

  ##subset years that you want (2016-2021), and create raster stack
  yrs <- as.character(unique(year(gps_data$date)))
  
  SE_s <- raster()
  
  for(i in 1:length(yrs)){
  SE_sub <- subset(SE, paste0("xmidS_SD_", yrs[i]))
  
  SE_s <- stack(SE_sub, SE_s)
  }

  ##crop to study area polygon
  SE <- crop(SE_s, st_transform(studyareapoly, projection(SE_s)))

# julian date of start of spring
r <- stack("Z:/MODIS_NDVI/Bischof_calculations/springStart.grd")
  
  ##subset by years you want
  r_s <- raster()
  
  for(i in 1:length(yrs)){
  r_sub <- subset(r, paste0("X", yrs[i]))
  
  r_s <- stack(r_sub, r_s)
  }
  
  ##crop to study area polygon
  r <- crop(r_s, st_transform(studyareapoly, projection(r)))
  
  ##set SE values >8 to NA and save raster object
  r[SE > 8] <- NA
  saveRDS(r, "./Code output/Season Rasters/springstart_julday.rds")


# mean date of peak NDVI
r <- stack("Z:/MODIS_NDVI/Bischof_calculations/maxNDVIdate.grd")

  ##subset by years you want
  r_s <- raster()
  
  for(i in 1:length(yrs)){
  r_sub <- subset(r, paste0("X", yrs[i]))
  
  r_s <- stack(r_sub, r_s)
  }
  
  ##crop to study area polygon
  r <- crop(r_s, st_transform(studyareapoly, projection(r_s)))

  ##set SE values >8 to NA and save raster object
  r[SE > 8] <- NA
  saveRDS(r, "./Code output/Season Rasters/meanPeakNDVI_julday.rds")


# mean date of peak brown down
SE <- stack("Z:/MODIS_NDVI/Bischof_calculations/dlcparams.grd")

  ##subset years that you want (2016-2021), and create raster stack
  yrs <- as.character(unique(year(gps_data$date)))
  
  SE_s <- raster()
  
  for(i in 1:length(yrs)){
  SE_sub <- subset(SE, paste0("xmidA_SD_", yrs[i])) #use xmidA instead of xmidS used for SE above
  
  SE_s <- stack(SE_sub, SE_s)
  }

  ##crop to study area polygon
  SE <- crop(SE_s, st_transform(studyareapoly, projection(SE_s)))

 
r <- stack("Z:/MODIS_NDVI/Bischof_calculations/dlcparams.grd")

  ##subset by years you want
  r_s <- raster()
  
  for(i in 1:length(yrs)){
  r_sub <- subset(r, paste0("xmidA_", yrs[i]))
  
  r_s <- stack(r_sub, r_s)
  }
  
  #crop to study area polygon
  r <- crop(r_s, st_transform(studyareapoly, projection(r_s)))
  
  ##set SE values >15 to NA and save raster object
  r[SE > 15] <- NA
  saveRDS(r, "./Code output/Season Rasters/meanPeakBrownDown_julday.rds")
```

#Snow phenology

Use SNODAS snow depth rasters across study area to calculate snow on and snow 
off dates (date there is snow or no snow at pixel for 8 consecutive days). 
This uses source code (CalcSnowOn and CalcSnowOff) from Jerod.

Download_SNODAS function from Jerod stopped working 8/2022. Seems like an 
issue with how the rasters are stacked- maybe an rgeos thing?? Still needs
to be fixed.

```{r}
#download snow depth for August-December 2021 (not on GIS drive)
##***Need to fix- now getting error from download_SNODAS
# source("Z:/Snodas/code/apiToDF.R")
# source("Z:/Snodas/code/download_SNODAS.R")
# download_SNODAS(params = "SnowDepth", startDate = "2021-08-01", endDate = "2021-12-31",
#                 outDir = "~/UWyo/PhD project/HOTR-proj/Data/GIS data/")

#stack snodas snow depth layers for recently downloaded files
lf <- list.files("~/UWyo/PhD project/HOTR-proj/Data/GIS data/", pattern = glob2rx("SNODAS_2021*SnowDepth*"), full.names = T)

name_vec2 <- c(list.files("~/UWyo/PhD project/HOTR-proj/Data/GIS data/", pattern = glob2rx("SNODAS_2021*SnowDepth*")))

stk_n <- stack(lf)

name_vec2 <- name_vec2 %>% str_replace(".tif", "")

#rename layers
names(stk_n) <- name_vec2

#crop by study area
stk_n <- stk_n %>%
  crop(extent(st_transform(studyareapoly, projection(.))))
```


```{r}
# date the there is no snow on the ground for 7-14 days. 
# date that there is consistent snow on the ground
# bring in all the layers for gps data years (2016-2021)
yrs <- as.character(unique(year(gps_data$date)))

#create vector of pattern to subset by
sd_patt <- c()
for(i in 1:length(yrs)){sd_patt <- c(paste0("SNODAS_", yrs[i]), sd_patt)}


#stack snodas snow depth layers for all data years
stk <- raster()
name_vec <- c()

for(i in 1:length(sd_patt)){
  print(i)
lf <- list.files("Z:/Snodas/snowdepth/", pattern = sd_patt[i], full.names = TRUE)
name_vec <- c(name_vec, list.files("Z:/Snodas/snowdepth/", pattern = sd_patt[i]))

stk_lf <- stack(lf)

stk <- addLayer(stk, stk_lf)
}

rm(stk_lf)
name_vec <- name_vec %>% str_replace(".tif", "")

#crop by study area
stk <- stk %>%
  crop(extent(st_transform(studyareapoly, projection(.))))   # doing this speeds thing up a lot!!, takes a few hrs

#add new data
stk <- addLayer(stk, stk_n)
```


```{r}
#calculate snow on and snow off rasters

#load source code
source("./R Code/Source Code/CalcSnowOn.R")
source("./R Code/Source Code/CalcSnowOff.R")

#set NAs in snow data to 999 (changed snow on/off calculations in source code to ignore this value)
stk[is.na(stk)] <- 999

#rename snow depth layers
names(stk) <- c(name_vec, name_vec2)

#save snodas snowdepth 2016-2021
writeRaster(stk,  "./Code output/SNODAS_snowdepth_2016to2021_stack", bylayer = T, overwrite = T)

#load study area snodas snowdepth 2016-2021
fn <- list.files("./Code output/", pattern = glob2rx("SNODAS_snowdepth_2016to2021_stack*.gri"), full.names = T)
stk <- stack(fn)

#snow off julian date raster (subset raster stack by each year of data)
SnowOff_jul <- raster()

for(i in 1:length(yrs)){
  r_sub <- subset(stk, grep(yrs[i], names(stk)))
  sn_o <- calc(r_sub, CalcSnowOff)
  
  SnowOff_jul <- stack(SnowOff_jul, sn_o)
  }

names(SnowOff_jul) <- yrs

#convert from stack to brick with 6 layers for better saving/processing
snowOff_brick <- brick(SnowOff_jul, n1 = 6)

saveRDS(snowOff_brick, "./Code output/Season Rasters/snowOFF_julday.rds")

#snow on julian date raster
SnowOn_jul <- raster()

for(i in 1:length(yrs)){
  r_sub <- subset(stk, grep(yrs[i], names(stk)))
  sn_o <- calc(r_sub, CalcSnowOn)
  
  SnowOn_jul <- stack(SnowOn_jul, sn_o) 
}

names(SnowOn_jul) <- yrs

#convert from stack to brick with 6 layers for better saving/processing
snowOn_brick <- brick(SnowOn_jul, n1 = 6)

saveRDS(snowOn_brick, "./Code output/Season Rasters/snowON_julday.rds")

```

#Calculate study area season start/end dates
The season start/end for the study area per year were calculated as the date where
60% of the pixels in the study area had reached the following criteria:
  Spring green-up: start- snow off, end- max NDVI peak
  Summer: start - max NDVI peak, end- max peak brown down
  Fall: start- max peak brown down, end- snow on
  Winter: start- snow on, end- snow off
  

The 60% came from testing values from 50% to 90% of the pixels to determine which
cutoffs matched the seasonality of the Yellowstone ecosystem on average. Percentages
higher than 60% were biased by slow snow melt at highest elevations, with spring
start dates being set to after June.
  

```{r}
#load rasters

snowOn <- readRDS("./Code output/Season Rasters/snowON_julday.rds")
snowOff <- readRDS("./Code output/Season Rasters/snowOFF_julday.rds")
meanPeakNDVI <- readRDS("./Code output/Season Rasters/meanPeakNDVI_julday.rds")
meanPeakBD <- readRDS("./Code output/Season Rasters/meanPeakBrownDown_julday.rds")
```

```{r}
#visualize data
hist(snowOff)
hist(snowOn)
hist(meanPeakBD)
hist(meanPeakNDVI)
```

```{r}
#pull yrs of data to loop through
yrs <- unique(year(gps_data$date))

#loop through yrs, creating data frame of season starts/ends per year
seas_list <- list()

for(i in 1:length(yrs)){
spring_start <- quantile(subset(snowOff, grep(yrs[i], names(snowOff))), 
                         probs = 0.6, names = F, digits = 0)

spring_end <- quantile(subset(meanPeakNDVI, grep(yrs[i], names(meanPeakNDVI))), 
                         probs = 0.6, names = F, digits = 0)

summer_start <- spring_end + 1 #start of next season = previous season end + 1 day
summer_end <- quantile(subset(meanPeakBD, grep(yrs[i], names(meanPeakBD))), 
                         probs = 0.6, names = F, digits = 0)

fall_start <- summer_end + 1
fall_end <- quantile(subset(snowOn, grep(yrs[i], names(snowOn))), 
                         probs = 0.6, names = F, digits = 0)  

winter_start <- fall_end + 1
winter_end <- spring_start - 1

yr <- yrs[i]

seas_list[[i]] <- data.frame(yr, spring_start, spring_end, summer_start, summer_end,
                        fall_start, fall_end, winter_start, winter_end)
}

#combine list of dates as one data frame
seas_df <- bind_rows(seas_list)

#reformat wide df to long df
seas_ldf <- pivot_longer(seas_df, cols = -c("yr"), names_to = "seas_type", 
                         values_to = "seas_jul_day")

#add date column, converting julian day to date
seas_ldf <- seas_ldf %>%
  mutate(seas_date = as.Date(seas_jul_day, origin = paste0(yr-1, "-12-31")))

#save as csv
write.csv(seas_ldf, "./Code output/season_startend_ndvi-snowdepth_2016to2021.csv")
```



