---
title: "YNP ungulate network connections"
author: "Molly Caldwell"
date: "3/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#load packages
library(sf)
library(mapview)
library(tidyverse)
library(dplyr)
library(lubridate)
library(spatsoc)
library(data.table)
library(asnipe)
library(igraph)
library(circlize)
<<<<<<< HEAD
=======
library(ComplexHeatmap)
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c
library(ggplot2)
library(GGally)
library(network)
library(sna)
library(maps)
library(ggnetwork)
library(ggpubr)
library(raster)
library(rgdal)
library(ggmap)
library(geosphere)
<<<<<<< HEAD

# if (!requireNamespace("BiocManager", quietly=TRUE))
#     install.packages("BiocManager")
# BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)
=======
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c
```

#Prep data

```{r}
#load seasonal gps data- all spp, 2 hr fixes
data_all <- readRDS("./Data/GPS data/Cleaned/allspp_cleanedGPSdata_seasonal_3.2022.rds")
<<<<<<< HEAD

#subset 2018 spring
data_all <- data_all %>%
  filter(yr == 2018 & season == "spring")
=======
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c
```

```{r}
#define projection
proj <- st_crs(data_all)[[1]]

#split geometry into x y columns
data_xy <- data_all %>%
    mutate(x = unlist(purrr::map(data_all$geometry,1)),
           y = unlist(purrr::map(data_all$geometry,2)))

data_xy$geometry <- NULL
data_xy$date2 <- NULL

#convert data to data table for spatsoc
data_dt <- setDT(data_xy)

```

<<<<<<< HEAD
#Spatial network- by points (100 m, within 12 hours)
=======
#Spatial network- by points (100 m, within x hours)
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c

##Temporal grouping

```{r}
#temporal group with threshold of x hours
g_data <- group_times(DT = data_dt, datetime = 'date', threshold = "12 hours") #change hours depending on desired threshold
```

##Spatial grouping (points)

```{r}
#spatial group (by GPS points) with threshold 100 meters
group_pts(DT = g_data, threshold = 100, id = 'cid', 
                        coords = c('x', 'y'), timegroup = 'timegroup') #threshold units based on GPS data CRS
```

##Build observed network

```{r}
#Generate group by individual matrix
gbiMtrx_pts <- get_gbi(DT = g_data, group = 'group', id = 'cid')

#build observed social network from spatial grouping by points
net_pts <- get_network(gbiMtrx_pts, data_format = "GBI",
                       association_index = "SRI")
```

##Calculate observed network metrics and graph   

```{r}
#generate graph
g_net_pts <- graph.adjacency(net_pts, "undirected", diag = FALSE,
                             weighted = TRUE)


#metrics for all individuals
# net_pts_metrics <- data.table(centrality = evcent(g_net_pts)$vector,
#                               strength = graph.strength(g_net_pts),
#                               degree = degree(g_net_pts),
#                               ID = names(degree(g_net_pts)),
#                               yr = data_dt_ss[, unique(year)])

#plot graph

plot(g_net_pts)

g_net_pts %>%
  add_layout_(with_fr()) %>%
  plot()
```

```{r}
#chord diagram

#list each species cids
u_cid <- unique(data_all$cid)
bison <- u_cid[grepl('BI', u_cid)]
bighorn <- u_cid[grepl('BH', u_cid)]
pronghorn <- u_cid[grepl('PR', u_cid)]
deer <- u_cid[grepl('MD', u_cid)]
elk <- u_cid[grepl('EL', u_cid)]

#create df assigning color to each cid by species
color_df <- data.frame(cid = c(bison, bighorn, pronghorn, deer, elk), 
                 color = c(rep("orange", length.out = length(bison)), 
                           rep("red", length.out = length(bighorn)),
                           rep("purple", length.out = length(pronghorn)),
                           rep("green", length.out = length(deer)),
                           rep("blue", length.out = length(elk))))



grid.col = as.vector(color_df$color)

names(grid.col) <- color_df$cid

#chord diagram colored by species
species <- as.factor(c("bison", "bighorn", "deer", "elk", "pronghorn"))
  
##legend
  lgd_col = Legend(
labels = species,                 
title = "Species",                 
type = "point",                 
background = unique(grid.col))

<<<<<<< HEAD
#plot chord diagram
=======
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c
plot.new()
chordDiagram(net_pts, grid.col = grid.col, scale = TRUE, annotationTrack = "grid",
             annotationTrackHeight = c(0.03, 0.01))
legend("bottomright", bty = "n", legend = species, fill = c("orange", "red", "green",
                                              "blue", "purple"))
```

##Network graph over landscape map and species' connection type summary

```{r}
  
#create network object
nw_line <- network(net_pts, directed = TRUE)

#create geocoded node coordinates (based on centroid of all GPS points by cid)
  ##create df of average location for each cid 
cid_avg <- data_dt %>%
  group_by(cid) %>%
  summarise(lat = mean(x), long = mean(y))

  ##add coords to network object
nw_line %v% "lat" <- if_else(cid_avg$cid == network.vertex.names(nw_line), cid_avg$lat, 999)
nw_line %v% "lon" <- if_else(cid_avg$cid == network.vertex.names(nw_line), cid_avg$long, 999)

#add species to network object
nw_line %v% "species" <- if_else(grepl('BI', network.vertex.names(nw_line)), "bison",
                          if_else(grepl('BH', network.vertex.names(nw_line)), "bighorn",
                          if_else(grepl('MD', network.vertex.names(nw_line)), "deer",
                          if_else(grepl('PR', network.vertex.names(nw_line)), "pronghorn",
                          if_else(grepl('EL', network.vertex.names(nw_line)), "elk", "NA")))))

# compute degree centrality
nw_line %v% "degree" <- degree(nw_line, gmode = "digraph")
```

###Network map

```{r}
#create data frame of nodes and node locations
nodes_cid <- nw_line$val
nodes_cid_df <- do.call(rbind.data.frame, nodes_cid)

#set projection to gps data
proj <- st_crs(data_all)[[1]]

#reproject and set as spatial points object for mapping
nodes_cid_sp <- SpatialPointsDataFrame(nodes_cid_df, 
                                       coords = nodes_cid_df[,3:4], 
                                       proj4string = CRS(proj))


nodes_cid_tr <- spTransform(nodes_cid_sp, CRS("+init=epsg:4326"))
nodes_cid_tr <- as.data.frame(nodes_cid_tr)
nodes_cid_tr$lat <- NULL
nodes_cid_tr$lon <- NULL

nodes_cid_tr <- nodes_cid_tr %>%
  rename(lat = lat.1, lon = lon.1)

#create data frame of node connections and frequencies

##extract data of head and tails of node connections
lines_cid <- as.data.frame(nw_line)

##transform heads and tails to source and target
conn_cid <- lines_cid %>%
  rename(target = .tail, source = .head)

##summarize frequency of connections
conn_cid <- conn_cid %>%
  group_by(target, source) %>%
  dplyr::mutate(freq = 1)


#create basemap of study area
<<<<<<< HEAD
  #ynp <- get_googlemap(center = c(lon = -110.50, lat = 44.9), zoom = 10, 
                    # maptype = "hybrid")
  #^not working since google update with access key needed...

# Tried this workaround with basemaps, but it's not communicating with ggplot- needs to be fixed
#library(basemaps)
# 
# set_defaults(map_service = "osm", map_type = "topographic")
# 
# ext <- st_make_grid(st_bbox(data_all), n = 1) #create extent based on gps data
# 
# base_map <- basemap_ggplot(ext) + labs(x = "Longitude", y = "Latitude")
=======
  ynp <- get_googlemap(center = c(lon = -110.50, lat = 44.9), zoom = 10, 
                     maptype = "hybrid")
library(basemaps)

set_defaults(map_service = "osm", map_type = "topographic")

ext <- st_make_grid(st_bbox(data_all), n = 1) #create extent based on gps data

base_map <- basemap_ggplot(ext) + labs(x = "Longitude", y = "Latitude")

 #create arcs for node connections
# for(i in 1:nrow(conn_cid)){
#     node1 <- nodes_cid_tr[nodes_cid_tr$vertex.names == conn_cid[i,]$source,]
#     node2 <- nodes_cid_tr[nodes_cid_tr$vertex.names == conn_cid[i,]$target,]
#     
#     arc <- gcIntermediate(c(node1[1,]$lat, node1[1,]$lon), 
#                            c(node2[1,]$lat, node2[1,]$lon), 
#                            n=1000, addStartEnd=TRUE )
#     
# }
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c


#add lat and lon to cid connections to create arcs (x,y and xend,yend)
nodes_xy <- nodes_cid_tr %>%
  dplyr::select(vertex.names, lat, lon, species)

conn_cid_xy <- left_join(conn_cid, nodes_xy, by = c("source" = "vertex.names"))
conn_cid_xy <- left_join(conn_cid_xy, nodes_xy, by = c("target" = "vertex.names"))

#add metric for proportion connections same species/different spp
  ##identify whether each connection is same species or different
  conn_cid_xy <- conn_cid_xy %>%
    mutate(same_spp = if_else(species.x == species.y, 1, 0)) %>%
    mutate(diff_spp = if_else(species.x != species.y, 1, 0))

  ##create data frame with sum of same spp/diff spp connections per source cid
  conn_spp_summ <- conn_cid_xy %>%
    group_by(source) %>%
    summarize(same_spp = sum(same_spp), diff_spp = sum(diff_spp))
  
  ##add percent connections that are different species
  conn_spp_summ <- conn_spp_summ %>%
    mutate(perc_diff = (diff_spp/(same_spp + diff_spp)*100))
  
  ##join same/diff connection summary to node data
  nodes_cid_tr <- left_join(nodes_cid_tr, conn_spp_summ, 
                            by = c("vertex.names" = "source"))
  
  ##add 0s for same/different connections for individuals with 0 connections
  nodes_cid_tr <- nodes_cid_tr %>%
    mutate(perc_diff = if_else(is.na(perc_diff), 0, perc_diff)) %>%
    rename("Percent interspecific connections" = perc_diff)
  
<<<<<<< HEAD
#map  nodes and connection lines- add basemap once that's working again
  ggplot() + 
=======
#map basemap, nodes, and connection lines
ggplot() +
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c
  geom_curve(data = conn_cid_xy, aes(x = lat.x, y = lon.x, xend = lat.y, 
                                     yend = lon.y, color = species.x),
                      curvature = 0.1, angle = 90, alpha = 0.4, size = 1,
                      show.legend = FALSE) +
  geom_point(data = nodes_cid_tr, aes(x = lat, y = lon, fill = species,
                                      size = `Percent interspecific connections`), 
             colour="black", pch = 21, stroke = 1.2, alpha = 0.7) + 
  scale_color_manual(values = c("bison" = "darkgoldenrod3",
                     "bighorn" = "red2",
                     "deer" = "chartreuse3",
                     "elk" = "dodgerblue3",
                     "pronghorn" = "darkorchid4")) +
  scale_fill_manual(name = "Individual centroids", values = c("bison" = "darkgoldenrod3",
                     "bighorn" = "red2",
                     "deer" = "chartreuse3",
                     "elk" = "dodgerblue3",
                     "pronghorn" = "darkorchid4")) + coord_cartesian()

```

###Species connection type summary and graphs

```{r}
#summary of species connections
  ##add 0s for same/different connections for individuals with 0 connections
  nodes_cid_tr <- nodes_cid_tr %>%
    mutate(same_spp = if_else(is.na(same_spp), 0, same_spp)) %>%
    mutate(diff_spp = if_else(is.na(diff_spp), 0, diff_spp)) %>%
    #rename(perc_diff = `Percent interspecific connections`) %>%
  mutate(tot_conn = same_spp + diff_spp)

#summary table- means by species
nodes_spp_summ <- nodes_cid_tr %>%
  group_by(species) %>%
  summarize(mean_tot_conn = mean(tot_conn), sd_tot_conn = sd(tot_conn), 
            mean_same_spp_conn = mean(same_spp), 
            sd_same_spp_conn = sd(same_spp),
            mean_diff_spp_conn = mean(diff_spp), 
            sd_diff_spp_conn = sd(diff_spp), .groups = "keep")
```

```{r}
library(scales)
#graph summaries spp connections
tot <- ggplot(nodes_spp_summ, aes(x = species, y = mean_tot_conn, color = species)) +
  geom_errorbar(aes(ymin = (mean_tot_conn - sd_tot_conn), 
                    ymax = (mean_tot_conn + sd_tot_conn)),
                width = 0.2) +
  scale_y_continuous(limits = c(0,55), oob = rescale_none) +
  geom_point() +
  xlab("Species") +
  ylab("Mean number of connections") +
  ggtitle("Total connections") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# same <- ggplot(nodes_spp_summ, aes(x = species, y = mean_same_spp_conn, color = species)) +
#   geom_errorbar(aes(ymin = (mean_same_spp_conn - sd_same_spp_conn), 
#                     ymax = (mean_same_spp_conn + sd_same_spp_conn)),
#                 width = 0.2) +
#   scale_y_continuous(limits = c(0,55), oob = rescale_none) +
#   geom_point() +
#   xlab("Species") +
#   ylab("") +
#   ggtitle("Intraspecies connections") +
#   theme(legend.position = "none") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# 
# diff <- ggplot(nodes_spp_summ, aes(x = species, y = mean_diff_spp_conn, color = species)) +
#   geom_errorbar(aes(ymin = (mean_diff_spp_conn - sd_diff_spp_conn), 
#                     ymax = (mean_diff_spp_conn + sd_diff_spp_conn)),
#                 width = 0.2) +
#   scale_y_continuous(limits = c(0,55), oob = rescale_none) +
#   geom_point() +
#   xlab("Species") +
#   ylab("") +
#   ggtitle("Interspecies connections") +
#   theme(legend.position = "none") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# 
# #plots on same page
# ggarrange(tot, same, diff, ncol = 3)
```

```{r}
#proportion intra and interspecies connections bar graph
  ##create data of proportions connections
  nodes_prop <- nodes_cid_tr %>%
    mutate(prop_same = same_spp/tot_conn) %>%
    mutate(prop_diff = diff_spp/tot_conn) %>%
    mutate(prop_same = if_else(is.na(prop_same), 0, prop_same)) %>%
    mutate(prop_diff = if_else(is.na(prop_diff), 0, prop_diff))

  ##average proportions by spp
  nodes_prop_mean <- nodes_prop %>%
    group_by(species) %>%
    summarize(intraspecies = mean(prop_same),
              interspecies = mean(prop_diff),
              .groups = "keep")
  
  ##sd proportions by spp
  nodes_prop_sd <- nodes_prop %>%
    group_by(species) %>%
    summarize(intraspecies = sd(prop_same),
              interspecies = sd(prop_diff),
              .groups = "keep")
  
  ##stack data for graphing
  prop_mean_stack <- melt(nodes_prop_mean, id.var = c("species"), 
                          variable.name = "connection_type",
                          value.name = "mean_proportion")
  prop_sd_stack <- melt(nodes_prop_sd, id.var = c("species"),
                        variable.name = "connection_type",
                        value.name = "sd_proportion")
  
  ##join mean and sd data
  prop_conn <- left_join(prop_mean_stack, prop_sd_stack, 
                         by = c("species", "connection_type"))
  
#graph
ggplot(prop_conn, aes(fill = connection_type, y = mean_proportion, x = species)) +
    geom_bar(position = "dodge", stat = "identity", alpha = 0.5) +
    geom_errorbar(aes(ymin = mean_proportion - sd_proportion, 
                      ymax = mean_proportion + sd_proportion), width=0.2,
                  colour="orange", position = position_dodge(.9)) +
    scale_y_continuous(limits = c(0,1), oob = rescale_none) +
  xlab("Species") +
  ylab("Mean proportion of total connections") +
  ggtitle("Proportion of intra- and interspecies connections") +
    scale_fill_discrete(name = "Connection type")
```

```{r}
#proportion interspecies connections with species colored graph
spp_conn <- conn_cid_xy %>%
  group_by(species.x, species.y) %>%
  summarize(freq_spp = n(), .groups = "keep") %>%
  mutate(conn_type = if_else(species.x == species.y, "Same", "Diff.")) %>%
  mutate(species.x = if_else(species.x == "bighorn", "Bighorn",
                   if_else(species.x == "bison", "Bison",
                   if_else(species.x == "pronghorn", "Pronghorn",
                   if_else(species.x == "elk", "Elk",
                   if_else(species.x == "deer", "Mule deer", "NA")))))) %>%
  mutate(species.y = if_else(species.y == "bighorn", "Bighorn",
                   if_else(species.y == "bison", "Bison",
                   if_else(species.y == "pronghorn", "Pronghorn",
                   if_else(species.y == "elk", "Elk",
                   if_else(species.y == "deer", "Mule deer", "NA")))))) %>%
  mutate(conn_type = factor(conn_type, levels = c("Same", "Diff."))) %>%
  rename("Species" = species.y) %>%
  ungroup() %>%
  group_by(species.x) %>%
  mutate(prop_conn = freq_spp/sum(freq_spp)) 

p1 <- ggplot(spp_conn, aes(x = conn_type, y = prop_conn, fill = Species)) +
  geom_col(position = "stack") +
  facet_grid(~species.x) +
  scale_fill_manual(values = c("#D55E00", "#E69F00", "#0072B2", "#009E73", "#CC79A7")) +
  theme(text = element_text(size = 16)) +
  xlab("Connection type") +
  ylab("Proportion connections") +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_y_continuous(limits = c(0,1), oob = rescale_none) 


g1 <- ggplot_gtable(ggplot_build(p1))
stripr <- which(grepl('strip-t', g1$layout$name))
fills <- c("#D55E00", "#E69F00", "#0072B2", "#009E73", "#CC79A7")
k <- 1
for (i in stripr) {
j <- which(grepl('rect', g1$grobs[[i]]$grobs[[1]]$childrenOrder))
g1$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
k <- k+1
}
grid.draw(g1)
```



<<<<<<< HEAD
####1hr- same code as above, 1hr time threshold
#Spatial network- by points (100 m, within 1 hours)
=======


####1hr
#Spatial network- by points (100 m, within x hours)
>>>>>>> 5cda1eb6d56219fc13b7bc11cede02ed0a3db28c

##Temporal grouping

```{r}
#temporal group with threshold of x hours
g_data1 <- group_times(DT = data_dt, datetime = 'date', threshold = "1 hour") #change hours depending on desired threshold
```

##Spatial grouping (points)

```{r}
#spatial group (by GPS points) with threshold 100 meters
group_pts(DT = g_data1, threshold = 100, id = 'cid', 
                        coords = c('x', 'y'), timegroup = 'timegroup') #threshold units based on GPS data CRS
```

##Build observed network

```{r}
#Generate group by individual matrix
gbiMtrx_pts1 <- get_gbi(DT = g_data1, group = 'group', id = 'cid')

#build observed social network from spatial grouping by points
net_pts1 <- get_network(gbiMtrx_pts1, data_format = "GBI",
                       association_index = "SRI")
```

##Calculate observed network metrics and graph   

```{r}
#generate graph
g_net_pts1 <- graph.adjacency(net_pts, "undirected", diag = FALSE,
                             weighted = TRUE)


#metrics for all individuals
# net_pts_metrics <- data.table(centrality = evcent(g_net_pts)$vector,
#                               strength = graph.strength(g_net_pts),
#                               degree = degree(g_net_pts),
#                               ID = names(degree(g_net_pts)),
#                               yr = data_dt_ss[, unique(year)])

#plot graph

plot(g_net_pts)

g_net_pts %>%
  add_layout_(with_fr()) %>%
  plot()
```

```{r}
#chord diagram

#list each species cids
u_cid <- unique(data_all$cid)
bison <- u_cid[grepl('BI', u_cid)]
bighorn <- u_cid[grepl('BH', u_cid)]
pronghorn <- u_cid[grepl('PR', u_cid)]
deer <- u_cid[grepl('MD', u_cid)]
elk <- u_cid[grepl('EL', u_cid)]

#create df assigning color to each cid by species
color_df <- data.frame(cid = c(bison, bighorn, pronghorn, deer, elk), 
                 color = c(rep("orange", length.out = length(bison)), 
                           rep("red", length.out = length(bighorn)),
                           rep("purple", length.out = length(pronghorn)),
                           rep("green", length.out = length(deer)),
                           rep("blue", length.out = length(elk))))



grid.col = as.vector(color_df$color)

names(grid.col) <- color_df$cid

#chord diagram colored by species
species <- as.factor(c("bison", "bighorn", "deer", "elk", "pronghorn"))
  
##legend
  lgd_col = Legend(
labels = species,                 
title = "Species",                 
type = "point",                 
background = unique(grid.col))

plot.new()
chordDiagram(net_pts, grid.col = grid.col, scale = TRUE, annotationTrack = "grid",
             annotationTrackHeight = c(0.03, 0.01))
legend("bottomright", bty = "n", legend = species, fill = c("orange", "red", "green",
                                              "blue", "purple"))
```

##Network graph over landscape map and species' connection type summary

```{r}
  
#create network object
nw_line1 <- network(net_pts1, directed = TRUE)

#create geocoded node coordinates (based on centroid of all GPS points by cid)
  ##create df of average location for each cid 
cid_avg1 <- data_dt %>%
  group_by(cid) %>%
  summarise(lat = mean(x), long = mean(y))

  ##add coords to network object
nw_line1 %v% "lat" <- if_else(cid_avg1$cid == network.vertex.names(nw_line1), cid_avg1$lat, 999)
nw_line1 %v% "lon" <- if_else(cid_avg1$cid == network.vertex.names(nw_line1), cid_avg1$long, 999)

#add species to network object
nw_line1 %v% "species" <- if_else(grepl('BI', network.vertex.names(nw_line1)), "bison",
                          if_else(grepl('BH', network.vertex.names(nw_line1)), "bighorn",
                          if_else(grepl('MD', network.vertex.names(nw_line1)), "deer",
                          if_else(grepl('PR', network.vertex.names(nw_line1)), "pronghorn",
                          if_else(grepl('EL', network.vertex.names(nw_line1)), "elk", "NA")))))

# compute degree centrality
nw_line1 %v% "degree" <- degree(nw_line1, gmode = "digraph")
```

###Network map

```{r}
#create data frame of nodes and node locations
nodes_cid1 <- nw_line1$val
nodes_cid_df1 <- do.call(rbind.data.frame, nodes_cid1)

#set projection to gps data
proj <- st_crs(data_all)[[1]]

#reproject and set as spatial points object for mapping
nodes_cid_sp1 <- SpatialPointsDataFrame(nodes_cid_df1, 
                                       coords = nodes_cid_df1[,3:4], 
                                       proj4string = CRS(proj))


nodes_cid_tr1 <- spTransform(nodes_cid_sp1, CRS("+init=epsg:4326"))
nodes_cid_tr1 <- as.data.frame(nodes_cid_tr1)
nodes_cid_tr1$lat <- NULL
nodes_cid_tr1$lon <- NULL

nodes_cid_tr1 <- nodes_cid_tr1 %>%
  rename(lat = lat.1, lon = lon.1)

#create data frame of node connections and frequencies

##extract data of head and tails of node connections
lines_cid1 <- as.data.frame(nw_line1)

##transform heads and tails to source and target
conn_cid1 <- lines_cid1 %>%
  rename(target = .tail, source = .head)

##summarize frequency of connections
conn_cid1 <- conn_cid1 %>%
  group_by(target, source) %>%
  dplyr::mutate(freq = 1)


#create basemap of study area
  ynp <- get_googlemap(center = c(lon = -110.50, lat = 44.9), zoom = 10, 
                     maptype = "hybrid")
library(basemaps)

set_defaults(map_service = "osm", map_type = "topographic")

ext <- st_make_grid(st_bbox(data_all), n = 1) #create extent based on gps data

base_map <- basemap_ggplot(ext) + labs(x = "Longitude", y = "Latitude")

 #create arcs for node connections
# for(i in 1:nrow(conn_cid)){
#     node1 <- nodes_cid_tr[nodes_cid_tr$vertex.names == conn_cid[i,]$source,]
#     node2 <- nodes_cid_tr[nodes_cid_tr$vertex.names == conn_cid[i,]$target,]
#     
#     arc <- gcIntermediate(c(node1[1,]$lat, node1[1,]$lon), 
#                            c(node2[1,]$lat, node2[1,]$lon), 
#                            n=1000, addStartEnd=TRUE )
#     
# }


#add lat and lon to cid connections to create arcs (x,y and xend,yend)
nodes_xy1 <- nodes_cid_tr1 %>%
  dplyr::select(vertex.names, lat, lon, species)

conn_cid_xy1 <- left_join(conn_cid1, nodes_xy1, by = c("source" = "vertex.names"))
conn_cid_xy1 <- left_join(conn_cid_xy1, nodes_xy1, by = c("target" = "vertex.names"))

#add metric for proportion connections same species/different spp
  ##identify whether each connection is same species or different
  conn_cid_xy1 <- conn_cid_xy1 %>%
    mutate(same_spp = if_else(species.x == species.y, 1, 0)) %>%
    mutate(diff_spp = if_else(species.x != species.y, 1, 0))

  ##create data frame with sum of same spp/diff spp connections per source cid
  conn_spp_summ1 <- conn_cid_xy1 %>%
    group_by(source) %>%
    summarize(same_spp = sum(same_spp), diff_spp = sum(diff_spp))
  #save csv
  # write.csv(conn_spp_summ1, "./Code output/allspp_alldata_1hrnetcon_byindiv_4.25.22.csv")
  
  ##add percent connections that are different species
  conn_spp_summ1 <- conn_spp_summ1 %>%
    mutate(perc_diff = (diff_spp/(same_spp + diff_spp)*100))
  
  ##join same/diff connection summary to node data
  nodes_cid_tr1 <- left_join(nodes_cid_tr1, conn_spp_summ1, 
                            by = c("vertex.names" = "source"))
  
  ##add 0s for same/different connections for individuals with 0 connections
  nodes_cid_tr1 <- nodes_cid_tr1 %>%
    mutate(perc_diff = if_else(is.na(perc_diff), 0, perc_diff)) %>%
    rename("Percent interspecific connections" = perc_diff)
  
#map basemap, nodes, and connection lines
ggplot() +
  geom_curve(data = conn_cid_xy, aes(x = lat.x, y = lon.x, xend = lat.y, 
                                     yend = lon.y, color = species.x),
                      curvature = 0.1, angle = 90, alpha = 0.4, size = 1,
                      show.legend = FALSE) +
  geom_point(data = nodes_cid_tr, aes(x = lat, y = lon, fill = species,
                                      size = `Percent interspecific connections`), 
             colour="black", pch = 21, stroke = 1.2, alpha = 0.7) + 
  scale_color_manual(values = c("bison" = "darkgoldenrod3",
                     "bighorn" = "red2",
                     "deer" = "chartreuse3",
                     "elk" = "dodgerblue3",
                     "pronghorn" = "darkorchid4")) +
  scale_fill_manual(name = "Individual centroids", values = c("bison" = "darkgoldenrod3",
                     "bighorn" = "red2",
                     "deer" = "chartreuse3",
                     "elk" = "dodgerblue3",
                     "pronghorn" = "darkorchid4")) + coord_cartesian()

```

###Species connection type summary and graphs

```{r}
#summary of species connections
  ##add 0s for same/different connections for individuals with 0 connections
  nodes_cid_tr1 <- nodes_cid_tr1 %>%
    mutate(same_spp = if_else(is.na(same_spp), 0, same_spp)) %>%
    mutate(diff_spp = if_else(is.na(diff_spp), 0, diff_spp)) %>%
    #rename(perc_diff = `Percent interspecific connections`) %>%
  mutate(tot_conn = same_spp + diff_spp)

#summary table- means by species
nodes_spp_summ1 <- nodes_cid_tr1 %>%
  group_by(species) %>%
  summarize(mean_tot_conn = mean(tot_conn), sd_tot_conn = sd(tot_conn), 
            mean_same_spp_conn = mean(same_spp), 
            sd_same_spp_conn = sd(same_spp),
            mean_diff_spp_conn = mean(diff_spp), 
            sd_diff_spp_conn = sd(diff_spp), .groups = "keep")
```

```{r}
library(scales)
#graph summaries spp connections
tot <- ggplot(nodes_spp_summ, aes(x = species, y = mean_tot_conn, color = species)) +
  geom_errorbar(aes(ymin = (mean_tot_conn - sd_tot_conn), 
                    ymax = (mean_tot_conn + sd_tot_conn)),
                width = 0.2) +
  scale_y_continuous(limits = c(0,55), oob = rescale_none) +
  geom_point() +
  xlab("Species") +
  ylab("Mean number of connections") +
  ggtitle("Total connections") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# same <- ggplot(nodes_spp_summ, aes(x = species, y = mean_same_spp_conn, color = species)) +
#   geom_errorbar(aes(ymin = (mean_same_spp_conn - sd_same_spp_conn), 
#                     ymax = (mean_same_spp_conn + sd_same_spp_conn)),
#                 width = 0.2) +
#   scale_y_continuous(limits = c(0,55), oob = rescale_none) +
#   geom_point() +
#   xlab("Species") +
#   ylab("") +
#   ggtitle("Intraspecies connections") +
#   theme(legend.position = "none") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# 
# diff <- ggplot(nodes_spp_summ, aes(x = species, y = mean_diff_spp_conn, color = species)) +
#   geom_errorbar(aes(ymin = (mean_diff_spp_conn - sd_diff_spp_conn), 
#                     ymax = (mean_diff_spp_conn + sd_diff_spp_conn)),
#                 width = 0.2) +
#   scale_y_continuous(limits = c(0,55), oob = rescale_none) +
#   geom_point() +
#   xlab("Species") +
#   ylab("") +
#   ggtitle("Interspecies connections") +
#   theme(legend.position = "none") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# 
# #plots on same page
# ggarrange(tot, same, diff, ncol = 3)
```

```{r}
#proportion intra and interspecies connections bar graph
  ##create data of proportions connections
  nodes_prop1 <- nodes_cid_tr1 %>%
    mutate(prop_same = same_spp/tot_conn) %>%
    mutate(prop_diff = diff_spp/tot_conn) %>%
    mutate(prop_same = if_else(is.na(prop_same), 0, prop_same)) %>%
    mutate(prop_diff = if_else(is.na(prop_diff), 0, prop_diff))

  ##average proportions by spp
  nodes_prop_mean1 <- nodes_prop1 %>%
    group_by(species) %>%
    summarize(intraspecies = mean(prop_same),
              interspecies = mean(prop_diff),
              .groups = "keep")
  
  ##sd proportions by spp
  nodes_prop_sd1 <- nodes_prop1 %>%
    group_by(species) %>%
    summarize(intraspecies = sd(prop_same),
              interspecies = sd(prop_diff),
              .groups = "keep")
  
  ##stack data for graphing
  prop_mean_stack <- melt(nodes_prop_mean, id.var = c("species"), 
                          variable.name = "connection_type",
                          value.name = "mean_proportion")
  prop_sd_stack <- melt(nodes_prop_sd, id.var = c("species"),
                        variable.name = "connection_type",
                        value.name = "sd_proportion")
  
  ##join mean and sd data
  prop_conn <- left_join(prop_mean_stack, prop_sd_stack, 
                         by = c("species", "connection_type"))
  
#graph
ggplot(prop_conn, aes(fill = connection_type, y = mean_proportion, x = species)) +
    geom_bar(position = "dodge", stat = "identity", alpha = 0.5) +
    geom_errorbar(aes(ymin = mean_proportion - sd_proportion, 
                      ymax = mean_proportion + sd_proportion), width=0.2,
                  colour="orange", position = position_dodge(.9)) +
    scale_y_continuous(limits = c(0,1), oob = rescale_none) +
  xlab("Species") +
  ylab("Mean proportion of total connections") +
  ggtitle("Proportion of intra- and interspecies connections") +
    scale_fill_discrete(name = "Connection type")
```

```{r}
#proportion interspecies connections with species colored graph
spp_conn <- conn_cid_xy1 %>%
  group_by(species.x, species.y) %>%
  summarize(freq_spp = n(), .groups = "keep") %>%
  mutate(conn_type = if_else(species.x == species.y, "Same", "Diff.")) %>%
  mutate(species.x = if_else(species.x == "bighorn", "Bighorn",
                   if_else(species.x == "bison", "Bison",
                   if_else(species.x == "pronghorn", "Pronghorn",
                   if_else(species.x == "elk", "Elk",
                   if_else(species.x == "deer", "Mule deer", "NA")))))) %>%
  mutate(species.y = if_else(species.y == "bighorn", "Bighorn",
                   if_else(species.y == "bison", "Bison",
                   if_else(species.y == "pronghorn", "Pronghorn",
                   if_else(species.y == "elk", "Elk",
                   if_else(species.y == "deer", "Mule deer", "NA")))))) %>%
  mutate(conn_type = factor(conn_type, levels = c("Same", "Diff."))) %>%
  rename("Species" = species.y) %>%
  ungroup() %>%
  group_by(species.x) %>%
  mutate(prop_conn = freq_spp/sum(freq_spp)) 

p2 <- ggplot(spp_conn, aes(x = conn_type, y = prop_conn, fill = Species)) +
  geom_col(position = "stack") +
  facet_grid(~species.x) +
  scale_fill_manual(values = c("#D55E00", "#E69F00", "#0072B2", "#009E73", "#CC79A7")) +
  theme(text = element_text(size = 16)) +
  xlab("Connection type") +
  ylab("Proportion connections") +
  theme(axis.text.x = element_text(angle = 45)) +
  theme(strip.background = element_blank(),
  strip.text.x = element_blank()) #remove facet labels

#color facet labels by spp
# g2 <- ggplot_gtable(ggplot_build(p2))
# stripr <- which(grepl('strip-t', g2$layout$name))
# fills <- c("#D55E00", "#E69F00", "#0072B2", "#009E73", "#CC79A7")
# k <- 1
# for (i in stripr) {
# j <- which(grepl('rect', g2$grobs[[i]]$grobs[[1]]$childrenOrder))
# g2$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
# k <- k+1
# }
# grid.draw(g2)
```