---
title: "YNP Ungulate Seasonal Movement Parameter Stats"
author: "Molly Caldwell"
date: "3/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#load packages
library(adehabitatLT)
library(adehabitatHR)
library(parallel)
library(circular)
library(sf)
library(lubridate)
library(moveHMM)
library(purrr)
library(dplyr)
library(ggplot2)
library(scales)
library(ggpubr)
```

```{r}
#load all species cleaned GPS collar data with move parameters (with season variables, elk and pronghorn subsampled to 2 hr, and individuals with missing seasonal data removed)
data_all <- readRDS("./Data/GPS data/Cleaned/allspp_cleanedGPSdata_seasonal_moveparams_3.2022.rds")

#add date only column
data_all <- data_all %>%
  mutate(date_only = format(date, format = "%Y-%m-%d"))
```

#Step length and turn angle

```{r, error=FALSE}
#list vector of species names
spp <- unique(data_all$species)

#data frame for combining summary stats
df_summ_stats <- data.frame(species = "NA", season = "NA", mean_speed = "NA",
                            sd_speed = "NA", mu = "NA", rho = "NA")

#list data by season
data_seas <- split(data_all, data_all$season)

#empty list to save species/season data
move_data_list <- list()

#Movement parameter graphs and summary data looped by season and species
for(i in 1:length(data_seas)){

  #select data from each season
data <- data_seas[[i]]

data <- data[order(data$id_yr_seas, data$date),] #order database first

#create season variable
season_select <- data$season[[i]]

#create data list by species to loop functions
data_list <- split(data, data$species)



  for(j in 1:length(data_list)){
  
  #write histograms of turn angle and step length to pngs
  # png(paste0("./Movement Parameters_Random Walks/Graphs/speed and turn angle 2021/", 
  #            spp[[j]], "_", data_list[[j]]$season[[i]], "_", 
  #            "turn_angle_hist", ".png"), width = 350, height = 350)
  # hist(data_list[[j]]$rel.angle, main = paste(spp[j], "turn angle", 
  #                                             season_select, sep = " "), 
  #      xlab = "turn angle (deg)")#this is your turning angle distribution
  # dev.off()
  # 
  # png(paste0("./Movement Parameters_Random Walks/Graphs/speed and turn angle 2021/", 
  #            spp[j], "_", data_list[[j]]$season[[i]], "_", 
  #            "speed_hist", ".png"), width = 350, height = 350)
  # hist((data_list[[j]]$speed), main = paste(spp[j], "speed", 
  #                                                     season_select, sep = " "), 
  #      xlab = "speed (km/hr)")   #this is your speed distribution
  # dev.off()
  
  #create list of mean and sd speed (km/hr) per species
  mean_Sp <- mean((data_list[[j]]$speed), na.rm = TRUE)
  sd_Sp <- sd((data_list[[j]]$speed), na.rm = TRUE)
  
  
  #create a list of mu and rho turn angle per species
  mu_rho_angle <- mle.wrappedcauchy(data_list[[j]]$rel.angle)
  
  #create data frame of combined movement parameter summaries
  new_df <- data.frame(species = spp[j], season = data_list[[j]]$season[[i]], 
                       mean_speed = mean_Sp, sd_speed = sd_Sp,
                       mu = mu_rho_angle$mu, rho = mu_rho_angle$rho)
  
  df_summ_stats <- rbind(df_summ_stats, new_df)
  
  }
move_data_list[[i]] <- data_list
}


#remove first "NA" row of summary stats data frame
df_summ_stats <- df_summ_stats %>% filter(species != "NA")
```

```{r}
#graph summary stats of speed per species and season
df_summ_stats <- df_summ_stats %>%
  mutate(mean_speed = as.numeric(mean_speed)) %>%
  mutate(sd_speed = as.numeric(sd_speed)) %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

ps <- ggplot(df_summ_stats, aes(x = species, y = mean_speed, color = species)) +
  geom_errorbar(aes(ymin = max((mean_speed - sd_speed), 0), ymax = (mean_speed + sd_speed)),
                width = 0.2) +
  geom_point() +
  scale_y_continuous(limits = c(0,0.2), oob = rescale_none) +
  facet_grid(cols = vars(season)) +
  xlab("Species") +
  ylab("Mean Speed (km/hr)") +
  ggtitle("Average speed") +
  theme(legend.position = "none") 

 ps + ggpubr::rotate()
 
 #save csv
 write.csv(df_summ_stats, 
           "./Code output/Species move metrics/spp_speed_turnangle_byseas.csv")
```

```{r}
#graph summary stats of turn angle per species and season
df_summ_stats <- df_summ_stats %>%
  mutate(mu = as.numeric(mu)) %>%
  mutate(rho = as.numeric(rho)) %>%
  mutate(mu_deg = mu * (180/pi)) %>%
  mutate(rho_deg = rho * (180/pi)) %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

pt <- ggplot(df_summ_stats, aes(x = species, y = mu_deg, color = species)) +
  geom_errorbar(aes(ymin = (mu_deg - rho_deg), ymax = (mu_deg + rho_deg)),
                width = 0.2) +
  geom_point() +
  facet_grid(cols = vars(season)) +
  xlab("Species") +
  ylab("mu turn angle (degrees)") +
  ggtitle("mu turn angle (+/- rho) by species and season") +
  theme(legend.position = "none")

pt + ggpubr::rotate()
 
```

##Daily distance traveled (sum of step lengths per day)

```{r}
#add total daily distance to movement data list
for(i in 1:length(move_data_list)){
  
  for(j in 1:length(move_data_list[[i]])){
    move_data_list[[i]][[j]] <- move_data_list[[i]][[j]] %>%
      group_by(id_yr_seas, date_only) %>%
      mutate(day_dist = sum(dist, na.rm = TRUE)) #sum step distance per day to get daily distance (in meters)
    
  }
  
}
```

```{r}
#create summary data of mean and sd daily distance by individual and season
#data frame for combining summary stats
summ_day_dist_indiv <- data.frame(id_yr_seas = "NA", species = "NA", season = "NA", 
                                  mean_day_dist = "NA", sd_day_dist = "NA")

#calculate mean and sd daily distance looped by cid and season
for(i in 1:length(move_data_list)){
  for(j in 1:length(move_data_list[[i]])){
    
    summ_dist <- move_data_list[[i]][[j]] %>%
      group_by(id_yr_seas) %>%
      summarise(mean_day_dist = mean(day_dist), 
                sd_day_dist = sd(day_dist))
    
    species <- move_data_list[[i]][[j]]$species[1:nrow(summ_dist)]
    season <- move_data_list[[i]][[j]]$season[1:nrow(summ_dist)]
    
    all_var_df <- cbind(summ_dist, species, season)
    all_var_df$geometry <- NULL
    
  #add data to summary df
    summ_day_dist_indiv <- rbind(all_var_df, summ_day_dist_indiv)
    
  }
  
}

#remove "NA" row of summary data frame
summ_day_dist_indiv <- summ_day_dist_indiv %>% filter(species != "NA")

```

```{r}
#graph mean and sd daily distance by individual and season (convert distance to km)
summ_day_dist_indiv <- summ_day_dist_indiv %>%
  mutate(mean_day_dist = as.numeric(mean_day_dist)/1000) %>%
  mutate(sd_day_dist = as.numeric(sd_day_dist)/1000) %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save csv
write.csv(summ_day_dist_indiv, 
          "./Code output/Species move metrics/indiv_dailydist_byseas.csv")

ggplot(summ_day_dist_indiv, aes(x = id_yr_seas, y = mean_day_dist, color = species)) +
  geom_errorbar(aes(ymin = (mean_day_dist - sd_day_dist), 
                    ymax = (mean_day_dist + sd_day_dist)),
                width = 0.2) +
  geom_point() +
  facet_grid(rows = vars(season)) +
  xlab("Collar ID") +
  ylab("Mean daily distance (km)") +
  ggtitle("Mean daily distance (km) by ID and season") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

```{r}
#create summary data of mean and sd daily distance by species and season
#data frame for combining summary stats
summ_day_dist_spp <- data.frame(species = "NA", season = "NA", 
                                  mean_day_dist = "NA", sd_day_dist = "NA")

#calculate mean and sd daily distance looped by cid and season
for(i in 1:length(move_data_list)){
  for(j in 1:length(move_data_list[[i]])){
    
    summ_dist <- move_data_list[[i]][[j]] %>%
      ungroup() %>%
      summarise(mean_day_dist = mean(day_dist), 
                sd_day_dist = sd(day_dist))
    
    species <- move_data_list[[i]][[j]]$species[1:nrow(summ_dist)]
    season <- move_data_list[[i]][[j]]$season[1:nrow(summ_dist)]
    
    all_var_df <- cbind(summ_dist, species, season)
    all_var_df$geometry <- NULL
    
  #add data to summary df
    summ_day_dist_spp <- rbind(all_var_df, summ_day_dist_spp)
    
  }
  
}

#remove "NA" row of summary data frame
summ_day_dist_spp <- summ_day_dist_spp %>% filter(species != "NA")
```

```{r}
#graph mean and sd daily distance by species and season (convert daily dist to km)
summ_day_dist_spp <- summ_day_dist_spp %>%
  mutate(mean_day_dist = as.numeric(mean_day_dist)/1000) %>%
  mutate(sd_day_dist = as.numeric(sd_day_dist)/1000) %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save csv
write.csv(summ_day_dist_spp, 
          "./Code output/Species move metrics/spp_dailydist_byseas.csv")

pdd <- ggplot(summ_day_dist_spp, aes(x = species, y = mean_day_dist, color = species)) +
  geom_errorbar(aes(ymin = (mean_day_dist - sd_day_dist), 
                     ymax = (mean_day_dist + sd_day_dist)),
                width = 0.2) +
  geom_point() +
  scale_y_continuous(limits = c(0,8), oob = rescale_none) +
  facet_grid(rows = vars(season)) +
  xlab("Species") +
  ylab("Mean daily distance (km)") +
  ggtitle("Mean daily distance") +
  theme(legend.position = "none")
  
  pdd + ggpubr::rotate()
```

##Seasonal distance (sum of daily distance per individual by season)

```{r}
#create summary data of mean and sd seasonal distance by individual and season
#data frame for combining summary stats
summ_seas_dist_indiv <- data.frame(id_yr_seas = "NA", species = "NA", season = "NA",
                                   seas_dist = "NA")

#calculate seasonal distance looped by cid and season
for(i in 1:length(move_data_list)){
  for(j in 1:length(move_data_list[[i]])){
    
    summ_dist <- move_data_list[[i]][[j]] %>%
      group_by(id_yr_seas) %>%
      summarise(seas_dist = sum(day_dist))
    
    species <- move_data_list[[i]][[j]]$species[1:nrow(summ_dist)]
    season <- move_data_list[[i]][[j]]$season[1:nrow(summ_dist)]
    
    all_var_df <- cbind(summ_dist, species, season)
    all_var_df$geometry <- NULL
    
  #add data to summary df
    summ_seas_dist_indiv <- rbind(all_var_df, summ_seas_dist_indiv)
    
  }
  
}

#remove "NA" row of summary data frame
summ_seas_dist_indiv <- summ_seas_dist_indiv %>% filter(species != "NA")

#summarize by species
summ_seas_dist_spp <- summ_seas_dist_indiv %>%
  mutate(seas_dist = as.numeric(seas_dist)/1000) %>% #convert to km
  group_by(species, season) %>%
  summarize(mean_seas_dist = mean(seas_dist), sd_seas_dist = sd(seas_dist),
            .groups = "keep") %>%
    mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save csv
write.csv(summ_seas_dist_spp, 
          "./Code output/Species move metrics/meanseas_dist_spp_byseas.csv")
```

```{r}
#graph mean and sd seasonal distance by species and season 
psd <- ggplot(summ_seas_dist_spp, aes(x = species, y = mean_seas_dist, color = species)) +
  geom_errorbar(aes(ymin = (mean_seas_dist - sd_seas_dist), 
                     ymax = (mean_seas_dist + sd_seas_dist)),
                width = 0.2) +
  geom_point() +
  scale_y_continuous(limits = c(0,6000), oob = rescale_none) +
  facet_grid(rows = vars(season)) +
  xlab("Species") +
  ylab("Mean seasonal distance (km)") +
  ggtitle("Mean seasonal distance") +
  theme(legend.position = "none")
  
  psd + ggpubr::rotate()
  

```

```{r}
#graph of only spring for tws wy poster 2022
  summ_seas_dist_spring <- summ_seas_dist_spp %>%
  filter(season == "spring") %>%
  mutate(species = if_else(species == "bighorn", "Bighorn",
                   if_else(species == "bison", "Bison",
                   if_else(species == "pronghorn", "Pronghorn",
                   if_else(species == "elk", "Elk",
                   if_else(species == "deer", "Mule deer", "NA"))))))

psd <- ggplot(summ_seas_dist_spring, aes(x = species, y = mean_seas_dist, color = species)) +
  geom_errorbar(aes(ymin = (mean_seas_dist - sd_seas_dist), 
                     ymax = (mean_seas_dist + sd_seas_dist)),
                width = 0.2, size = 2) +
  geom_point(size = 5) +
  scale_color_manual(values = c("#D55E00", "#E69F00", "#0072B2", "#009E73", "#CC79A7")) +
  scale_y_continuous(limits = c(0,6000), oob = rescale_none) +
  xlab("") +
  ylab("Mean distance traveled (km)") +
  theme(legend.position = "none") +
  theme(text = element_text(size = 20))
  
  psd + ggpubr::rotate()
```


#Maximum daily displacement
Defined as maximum distance between any two points during the same day by CID

```{r}
##create data frame to store max daily distance info per individual per season
max_day_disp_indiv <- data.frame(id_yr_seas = "NA", species = "NA", season = "NA",
                                date = "NA", max_disp_day = "NA")

##setup cluster to run on multiple cores
#set number of cores to use (1 less than you have)
no_cores <- detectCores()-1

# Setup cluster
clust <- makeCluster(no_cores) 
# export the objects you need for your calculations from your environment to each node's environment
clusterExport(clust, varlist=c("max_day_disp_indiv", "move_data_list"))


##loop by species, season, individual, and day - takes a few hrs
  for(i in 1:length(move_data_list)){
    print(i)
    for(j in 1:length(move_data_list[[i]])){
      u_cid <- unique(move_data_list[[i]][[j]]$id_yr_seas) #pull list of unique id_yr_seas
      
      for(k in 1:length(u_cid)){
        print(u_cid[[k]])
        dt <- move_data_list[[i]][[j]]
        dt_cid <- dt %>% filter(id_yr_seas == u_cid[[k]]) #create df of each cid
        
        
        u_day <- unique(dt_cid$date_only)
        
        for(l in 1:length(u_day)){
          dt_day <- dt_cid %>% filter(date_only == u_day[[l]]) #create df of each day
          
        #calculate distance between every point each day
        dist_mt <- st_distance(dt_day, dt_day)
        
        #extract max distance
        max_disp_seas <- max(dist_mt)
        
        #extract cid, season, species for data frame
        id_yr_seas <- u_cid[[k]]
        species <- move_data_list[[i]][[j]]$species[[1]]
        season <- move_data_list[[i]][[j]]$season[[1]]
        date <- u_day[[l]]
        
        #add to existing data frame
        df_md <- cbind(id_yr_seas, species, season, date, max_disp_seas)
        df_md$geometry <- NULL
        
        max_day_disp_indiv <- rbind(df_md, max_day_disp_indiv)
        
        }
      }
    }
  }

stopCluster(clust)   #stop the parallelization framework

#remove "NA" row of summary data frame
max_day_disp_indiv <- max_day_disp_indiv %>% filter(species != "NA")
```


```{r}
#graph max daily displacement by cid and season (convert dist to km)
max_day_disp_indiv <- max_day_disp_indiv %>%
  mutate(max_disp_day = as.numeric(max_disp_day)) %>%
  mutate(max_disp_day_km = as.numeric(max_disp_day/1000)) 

#summarize by mean and sd daily displacement by cid and season
summ_day_disp_indiv <- max_day_disp_indiv %>%
  group_by(id_yr_seas, season, species) %>%
  summarize(mean_disp = mean(max_disp_day_km), sd_disp = sd(max_disp_day_km),
            .groups = "keep") %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save csv
write.csv(summ_day_disp_indiv, "./Code output/Species move metrics/avgdaydisp_indiv_byseas.csv")

#graph
ggplot(summ_day_disp_indiv, aes(x = id_yr_seas, y = mean_disp, color = species)) +
  geom_errorbar(aes(ymin = (mean_disp - sd_disp), 
                    ymax = (mean_disp + sd_disp)),
                width = 0.2) +
  geom_point() +
  facet_grid(rows = vars(season)) +
  xlab("Collar ID") +
  ylab("Mean daily displacement (km)") +
  ggtitle("Mean daily displacement (+/- stand. dev.) by collar ID and season") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

```{r}
#graph max daily displacement by species and season (convert dist to km)

#summarize by mean and sd daily displacement by species and season
summ_day_disp_spp <- max_day_disp_indiv %>%
  group_by(species, season) %>%
  summarize(mean_disp = mean(max_disp_day_km), sd_disp = sd(max_disp_day_km),
            .groups = "keep") %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save csv
write.csv(summ_day_disp_spp, "./Code output/Species move metrics/spp_daydispavg_byseas.csv")

#graphp
pdd <- ggplot(summ_day_disp_spp, aes(x = species, y = mean_disp, color = species)) +
  geom_errorbar(aes(ymin = (mean_disp - sd_disp), 
                    ymax = (mean_disp + sd_disp)),
                width = 0.2) +
  geom_point() +
  facet_grid(rows = vars(season)) +
  xlab("Species") +
  ylab("Mean daily displacement (km)") +
  ggtitle("Mean daily displacement") +
  theme(legend.position = "none") 

pdd + ggpubr::rotate()
```


#Maximum seasonal displacement
Defined as maximum distance between any two points by species during same season

```{r}
##create data frame to store max displacement per individual by season
max_disp_id_seas <- data.frame(id_yr_seas = "NA", species = "NA", season = "NA",
                                max_disp_seas = "NA")

##loop by species, individual, and season 
  for(i in 1:length(move_data_list)){
    
    for(j in 1:length(move_data_list[[i]])){
      u_id_yr_seas <- unique(move_data_list[[i]][[j]]$id_yr_seas) #pull list of unique id_yr_seass
      
      for(k in 1:length(u_id_yr_seas)){
        dt <- move_data_list[[i]][[j]]
        dt_id_yr_seas <- dt %>% filter(id_yr_seas == u_id_yr_seas[[k]]) #create df of each id_yr_seas
        
        #calculate distance between every point
        dist_mt <- st_distance(dt_id_yr_seas, dt_id_yr_seas)
        
        #extract max distance
        max_disp_seas <- max(dist_mt)
        
        #extract id_yr_seas, season, species for data frame
        id_yr_seas <- u_id_yr_seas[[k]]
        species <- move_data_list[[i]][[j]]$species[[1]]
        season <- move_data_list[[i]][[j]]$season[[1]]
        
        #add to existing data frame
        df_md <- cbind(id_yr_seas, species, season, max_disp_seas)
        df_md$geometry <- NULL
        
        max_disp_id_seas <- rbind(df_md, max_disp_id_seas)
      }
    }
  }

#remove "NA" row of summary data frame
max_disp_id_seas <- max_disp_id_seas %>% filter(species != "NA")
```


```{r}
#graph max displacement by cid and season (convert dist to km)
max_disp_id_seas <- max_disp_id_seas %>%
  mutate(max_disp_seas = as.numeric(max_disp_seas)) %>%
  mutate(max_disp_seas_km = as.numeric(max_disp_seas/1000)) %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save as csv
write.csv(max_disp_id_seas, "./Code output/Species move metrics/indiv_disp_byseas.csv")

ggplot(max_disp_cid_seas, aes(x = cid, y = max_disp_seas_km, fill = species)) +
  geom_col() +
  facet_grid(rows = vars(season)) +
  xlab("Collar ID") +
  ylab("Maximum displacement distance (km)") +
  ggtitle("Maximum displacement by CID and season") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


```{r}
#graph max displacement by species and season (convert dist to km)
max_disp_spp_seas <- max_disp_id_seas %>% 
  group_by(species, season) %>%
  summarize(mean_disp = mean(max_disp_seas_km), sd_disp = sd(max_disp_seas_km),
            .groups = "keep") %>%
  mutate(season = factor(season, levels = c("summer", "fall", "winter", "spring")))

#save csv
write.csv(max_disp_spp_seas, "./Code output/Species move metrics/spp_avgseasdisp_byseas.csv")

pmds <- ggplot(max_disp_spp_seas, aes(x = species, y = mean_disp, color = species)) +
  geom_errorbar(aes(ymin = (mean_disp - sd_disp), 
                    ymax = (mean_disp + sd_disp)),
                width = 0.2) +
  geom_point() +
  facet_grid(rows = vars(season)) +
  xlab("Species") +
  ylab("Mean displacement (km)") +
  ggtitle("Mean seasonal displacement") +
  theme(legend.position = "none")

pmds + ggpubr::rotate()
```

```{r}
save.image("./YNP_ungulates_moveparams.RData")
```

