---
title: "YNP Ungulate Individual Basic Move Metrics: daily, weekly, monthly, seasonally, annually"
author: "Molly Caldwell"
date: "6/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#load packages
library(adehabitatLT)
library(adehabitatHR)
library(parallel)
library(circular)
library(sf)
library(lubridate)
library(moveHMM)
library(purrr)
library(dplyr)
```

```{r}
#load all species cleaned GPS collar data with move parameters (with season variables, elk and pronghorn subsampled to 2 hr, and individuals with missing seasonal data removed)
data_all <- readRDS("./Data/GPS data/Cleaned/allspp_cleanedGPSdata_seasonal_moveparams_3.2022.rds")

#add date only column
data_all <- data_all %>%
  mutate(date_only = format(date, format = "%Y-%m-%d"))

#remove extra date column
data_all$date2 <- NULL
```

#Daily basic move metrics
Average and standard deviation of speed, turn angle, and step length. Daily total distance and displacement (max distance between gps points) by individual.

```{r}
#avg and sd of speed and step length, sum step length, displacement
indiv_movemet_day <- data_all %>%
  group_by(id_yr_seas, date_only, species, season, yr) %>% #added in extra grouping columns for additional variables I want to keep
  summarize(avg_speed = mean(speed, na.rm = T), sd_speed = sd(speed, na.rm = T), 
            avg_steplength = mean(dist, na.rm = T), 
            sd_steplength = sd(dist, na.rm = T), 
            tot_dist = sum(dist, na.rm = T),
            max_displ = as.numeric(max(st_distance(t))), .groups = "keep")
         
   
```

```{r}
#loop avg and sd of turn angle by day and id_yr_seas b/c NAs throwing errors in summarize
ang_summ <- data.frame(id_yr_seas = "NA", date_only = "NA", mu_turnang = "NA",
                       rho_turnang = "NA")

u_id <- unique(data_all$id_yr_seas)

#run loop in parallel
#identify 1 less cores than you have
no_cores <- detectCores()-1
#setup cluster
clust <- makeCluster(no_cores)
#export needed objects to nodes' environments
clusterExport(clust, varlist = c("ang_summ", "u_id", "data_all"))

for(i in 1:length(u_id)){
  print(paste0("i: ", i))
  data_id <- data_all %>% filter(id_yr_seas == u_id[i])
  u_date <- unique(data_id$date_only)
  
  for(j in 1:length(u_date)){
    data_dt <- data_id %>% filter(date_only == u_date[j])
    
    if(sum(!is.na(data_dt$rel.ang)) > 1){
      mu_turnang <- as.numeric(mle.wrappedcauchy(data_dt$rel.angle)$mu)
      rho_turnang <- as.numeric(mle.wrappedcauchy(data_dt$rel.angle)$rho)
      id_yr_seas <- data_dt$id_yr_seas[[1]]
      date_only <- data_dt$date_only[[1]]
      
      df <- cbind(id_yr_seas, date_only, mu_turnang, rho_turnang)
      ang_summ <- rbind(df, ang_summ)
    }
  }
}

#stop parallelization framework
stopCluster(clust)

```

```{r}
#combine turn angle and move metric daily summaries
indiv_movemet_day <- left_join(indiv_movemet_day, ang_summ, 
                               by = c("id_yr_seas", "date_only"))

#remove na column
indiv_movemet_day <- indiv_movemet_day %>% filter(id_yr_seas != "NA")

#save as csv
write.csv(indiv_movemet_day, 
      "./Code output/individual_move_metrics/indiv_day_move_metrics_6_22.csv")
```


#Weekly (5 days) basic move metrics
Average and standard deviation of speed, turn angle, and step length. Weekly total distance and displacement (max distance between gps points) by individual.

```{r}
#create 5-day grouping variable per id_yr_seas
data_all <- data_all %>% group_by(id_yr_seas) %>%
  mutate(week = cut(as.Date(date), breaks = "5 days", include.lowest = T))
  
```


```{r}
#avg and sd of speed and step length, sum step length, displacement
indiv_movemet_wk <- data_all %>%
  group_by(id_yr_seas, week, species, season, yr) %>% #added in extra grouping columns for additional variables I want to keep
  summarize(avg_speed = mean(speed, na.rm = T), sd_speed = sd(speed, na.rm = T), 
            avg_steplength = mean(dist, na.rm = T), 
            sd_steplength = sd(dist, na.rm = T), 
            tot_dist = sum(dist, na.rm = T),
            max_displ = as.numeric(max(st_distance(geometry))), .groups = "keep")

indiv_movemet_wk$geometry <- NULL
         
```

```{r}
#loop avg and sd of turn angle b/c NAs throwing errors in summarize
ang_summ_wk <- data.frame(id_yr_seas = "NA", week = "NA", mu_turnang = "NA",
                       rho_turnang = "NA")

u_id <- unique(data_all$id_yr_seas)

#run loop in parallel
#identify 1 less cores than you have
no_cores <- detectCores()-1
#setup cluster
clust <- makeCluster(no_cores)
#export needed objects to nodes' environments
clusterExport(clust, varlist = c("ang_summ_wk", "u_id", "data_all"))

for(i in 1:length(u_id)){
  print(paste0("i: ", i))
  data_id <- data_all %>% filter(id_yr_seas == u_id[i])
  u_wk <- unique(data_id$week)
  
  for(j in 1:length(u_wk)){
    data_dt <- data_id %>% filter(week == u_wk[j])
    
    if(sum(!is.na(data_dt$rel.angle)) > 1){
      mu_turnang <- as.numeric(mle.wrappedcauchy(data_dt$rel.angle)$mu)
      rho_turnang <- as.numeric(mle.wrappedcauchy(data_dt$rel.angle)$rho)
      id_yr_seas <- data_dt$id_yr_seas[[1]]
      week <- as.character(data_dt$week[[1]])
      
      df <- cbind(id_yr_seas, week, mu_turnang, rho_turnang)
      ang_summ_wk <- rbind(df, ang_summ_wk)
    }
  }
}

#stop parallelization framework
stopCluster(clust)

```

```{r}
#combine turn angle and move metric daily summaries
indiv_movemet_wk <- left_join(indiv_movemet_wk, ang_summ_wk, 
                               by = c("id_yr_seas", "week"))


#save as csv
write.csv(indiv_movemet_wk, 
      "./Code output/individual_move_metrics/indiv_week_move_metrics_6_22.csv")
```

#Monthly basic move metrics
Average and standard deviation of speed, turn angle, and step length. Total distance and displacement (max distance between gps points) by individual.

```{r}
#create month grouping variable per id_yr_seas
data_all <- data_all %>% group_by(id_yr_seas) %>%
  mutate(month = cut(as.Date(date), breaks = "1 month", include.lowest = T))
  
```

```{r}
#avg and sd of speed and step length, sum step length, displacement
indiv_movemet_month <- data_all %>%
  group_by(id_yr_seas, month, species, season, yr) %>% #added in extra grouping columns for additional variables I want to keep
  summarize(avg_speed = mean(speed, na.rm = T), sd_speed = sd(speed, na.rm = T), 
            avg_steplength = mean(dist, na.rm = T), 
            sd_steplength = sd(dist, na.rm = T), 
            tot_dist = sum(dist, na.rm = T),
            max_displ = as.numeric(max(st_distance(geometry))), .groups = "keep")
         
indiv_movemet_month$geometry <- NULL
```

```{r}
#loop avg and sd of turn angle b/c NAs throwing errors in summarize
ang_summ_month <- data.frame(id_yr_seas = "NA", month = "NA", mu_turnang = "NA",
                       rho_turnang = "NA")

u_id <- unique(data_all$id_yr_seas)

#run loop in parallel
#identify 1 less cores than you have
no_cores <- detectCores()-1
#setup cluster
clust <- makeCluster(no_cores)
#export needed objects to nodes' environments
clusterExport(clust, varlist = c("ang_summ_month", "u_id", "data_all"))

for(i in 1:length(u_id)){
  print(paste0("i: ", i))
  data_id <- data_all %>% filter(id_yr_seas == u_id[i])
  u_month <- unique(data_id$month)
  
  for(j in 1:length(u_month)){
    data_dt <- data_id %>% filter(month == u_month[j])
    
    if(sum(!is.na(data_dt$rel.angle)) > 1){
      mu_turnang <- as.numeric(mle.wrappedcauchy(data_dt$rel.angle)$mu)
      rho_turnang <- as.numeric(mle.wrappedcauchy(data_dt$rel.angle)$rho)
      id_yr_seas <- data_dt$id_yr_seas[[1]]
      month <- as.character(data_dt$month[[1]])
      
      df <- cbind(id_yr_seas, month, mu_turnang, rho_turnang)
      ang_summ_month <- rbind(df, ang_summ_month)
    }
  }
}

#stop parallelization framework
stopCluster(clust)

```

```{r}
#combine turn angle and move metric daily summaries
indiv_movemet_month <- left_join(indiv_movemet_month, ang_summ_month, 
                               by = c("id_yr_seas", "month"))


#save as csv
write.csv(indiv_movemet_month, 
      "./Code output/individual_move_metrics/indiv_month_move_metrics_6_22.csv")
```


#Seasonal basic move metrics
Average and standard deviation of speed, turn angle, and step length. Total distance and displacement (max distance between gps points) by individual.


```{r}
#avg and sd of speed and step length, sum step length, displacement
indiv_movemet_seas <- data_all %>%
  group_by(id_yr_seas, species, season, yr) %>% #added in extra grouping columns for additional variables I want to keep
  summarize(avg_speed = mean(speed, na.rm = T), sd_speed = sd(speed, na.rm = T), 
            avg_steplength = mean(dist, na.rm = T), 
            sd_steplength = sd(dist, na.rm = T), 
            tot_dist = sum(dist, na.rm = T),
            mu_turnang = as.numeric(mle.wrappedcauchy(rel.angle)$mu),
            rho_turnang = as.numeric(mle.wrappedcauchy(rel.angle)$rho),
            max_displ = as.numeric(max(st_distance(geometry))), .groups = "keep")

indiv_movemet_seas$geometry <- NULL

#save as csv
write.csv(indiv_movemet_seas, 
      "./Code output/individual_move_metrics/indiv_seas_move_metrics_6_22.csv")
```

#Annual basic move metrics
Average and standard deviation of speed, turn angle, and step length. Total distance and displacement (max distance between gps points) by individual.


```{r}
#avg and sd of speed and step length, sum step length, displacement
indiv_movemet_ann <- data_all %>%
  group_by(cid, species, yr) %>% #added in extra grouping columns for additional variables I want to keep
  summarize(avg_speed = mean(speed, na.rm = T), sd_speed = sd(speed, na.rm = T), 
            avg_steplength = mean(dist, na.rm = T), 
            sd_steplength = sd(dist, na.rm = T), 
            tot_dist = sum(dist, na.rm = T),
            mu_turnang = as.numeric(mle.wrappedcauchy(rel.angle)$mu),
            rho_turnang = as.numeric(mle.wrappedcauchy(rel.angle)$rho),
            max_displ = as.numeric(max(st_distance(geometry))), .groups = "keep")
  
indiv_movemet_ann$geometry <- NULL     

#save as csv
write.csv(indiv_movemet_ann, 
      "./Code output/individual_move_metrics/indiv_annual_move_metrics_6_22.csv")
```


