---
title: "YNP ungulate state space models"
author: "Molly Caldwell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
# packages
library(adehabitatLT)
library(adehabitatHR)
library(parallel)
library(circular)
library(sf)
library(mapview)
library(lubridate)
library(tidyverse)
library(moveHMM)
#library(devtools) #use to update jerod's MoveTools package
#install_github("jmerkle1/MerkleLab-Code-Repository", subdir="MoveTools")
library(MoveTools)
```

#Prep Data

```{r}
#load green up data all species
data <- readRDS("./Data/GPS data/Cleaned/allspp_cleanedGPSdata_greenup(mar_to_jun).rds")

#create list data by species for loops
spp <- c("bighorn", "bison", "elk", "deer", "pronghorn")
data_list <- list()

for(i in 1:length(spp)){
  data_sp <- data %>% filter(species == spp[[i]])
  data_list[[i]] <- data_sp
}

rm(data_sp)

```

```{r}
## first pull out only the ID, date, and coordinates. you MUST name them exactly these things
## to use the functions in moveHMM
df_list <- list()

for(i in 1:length(data_list)){
  data <- data_list[[i]]
df <- data.frame(ID=data$id_yr_seas,
               Time=data$date,
               X=st_coordinates(data)[,1],
               Y=st_coordinates(data)[,2])

df_list[[i]] <- df
}

rm(data, df)

## this function preps the data to run a state space model
## put in the dataframe you just made, the coordinate type and the name of the coordinate columns
data.hmm_list <- list()

for(i in 1:length(df_list)){
data.hmm <- prepData(df_list[[i]], type="UTM", coordNames = c("X", "Y"))
data.hmm_list[[i]] <- data.hmm
}
```

```{r}
## this next step will first plot the movement of each of your individuals (colored by ID)
## and will then plot various movement parameters for each individual animal
plot(data.hmm_list[[1]], compact = TRUE)  

summary(data.hmm_list[[1]]) ##gives you summary information
```

```{r}
## if you have steps with step length zero, you have to account for that in the models
## by including a zero-mass starting parameter because the strictly positive distributions
## (i.e., gamma) are inadequate to use. You can either account for this by adding in a zero-mass
## parameter, or you can add a tiny amount of positive variation to the steps with length 0
## we are going to do that for similicity sake. For more information about how to deal with zero-
## inflation, you can read about it in the moveHMM vignette.

## first check to see if you have any steps with length 0
table(data.hmm_list[[1]]$step==0) ## 88 steps with step length 0
table(data.hmm_list[[2]]$step==0) ## 146 steps with step length 0
table(data.hmm_list[[3]]$step==0) ## 439 steps with step length 0
table(data.hmm_list[[4]]$step==0) ## 32 steps with step length 0
table(data.hmm_list[[5]]$step==0) ## 0 steps with step length 0

## add in some very small variation for steps with length 0
for(i in 1:length(data_list)){
data.hmm_list[[i]]$step <- ifelse(data.hmm_list[[i]]$step==0, 
                                  0+runif(1, 0, .001), data.hmm_list[[i]]$step)
}

## after you add in the variation, these should all be FALSE, if they are not, something is wrong
table(data.hmm_list[[5]]$step==0)
```

#State Space Models

```{r}
#starting parameters
## before we run models with different states, we need to set some starting parameters for the models to start from
## we can use the mean and distribution of our actual data to decide on these
## looking at the distribution of steps and turning angles will also let us decide what
## distribution to use
mean_step <- c()
sd_step <- c()
mean_angle <- c()
sd_angle <- c()

for(i in 1:length(data.hmm_list)){
hist(data.hmm_list[[i]]$step, main = spp[[i]])
hist(data.hmm_list[[i]]$angle, main = spp[[i]])

mean_step <- c(mean_step, mean(data.hmm_list[[i]]$step, na.rm=T))
sd_step <- c(sd_step, sd(data.hmm_list[[i]]$step, na.rm=T))

mean_angle <- c(mean_angle, mean(data.hmm_list[[i]]$angle, na.rm=T))
sd_angle <- c(sd_angle, sd(data.hmm_list[[i]]$angle, na.rm=T))
}


```

to fit all of the models, we will use the fitHMM function in moveHMM
for each of the models you run in the fitHMM function, you need to provide a starting point
for each state you are trying to identify.

the parameters depend on the distribution that you use, for the BHS data, we will use gamma and vm distribution
you can use gamma, Weibull, log-normal, or exponential for the step lengths
and vonMises or wrapped Cauchy for the angles
look at the vignette for more info on what parameters you should use for each distribution

if you get errors when running the models, and all data that you prepped looks correct
the errors likely are a result of the optimizer failing to converge
this happens if you set your initial parameters to something that doesn't quite make sense
i.e., if your average step length is 8 m and you set the mean to 8,000 it won't converge
so just mess around with the starting parameters until you get it to properly converge

okay, now we have a general idea of our movement data and know what distribution we want to use
we can start with the single-state model. For the single state model, choosing the starting parameters is
pretty easy - just put in the means and sd of the steps and angles

##Single State Model


```{r}
ss_model_list <- list()

for(i in 1:length(data.hmm_list)){
  print(paste0("i: ", i))
# give some starting parameter values:
mu0 <- mean_step[i] # step mean (need to set one for each state, so for single state, you only need one)
sigma0 <- sd_step[i] # step SD 
stepPar0 <- c(mu0,sigma0) ## combine the mean and SD into one object to fit in the model

angleMean0 <- mean_angle[i] # angle mean (need to set one for each state)
kappa0 <- sd_angle[i] # angle concentration
anglePar0 <- c(angleMean0,kappa0) ## combine into one object

## with verbose=2, you get information for each iteration, it tells you if you are getting close to the 
## solution. You can turn that off by setting verbose=0, or if you want to see only the first and last iteration, set
## verbose=1, The worse your starting parameters are, the longer it takes to find a solution (and more likely to fail)

singleStateModel <- fitHMM(data.hmm_list[[i]], nbStates = 1, 
                          stepPar0 = stepPar0, 
                          anglePar0 = anglePar0, verbose=2,
                          stepDist = "gamma", angleDist = "vm")

ss_model_list[[i]] <- singleStateModel

}

##can look at model output
ss_model_list[[1]] ## this gives you the mean and sd of your states (in this case a single state).

# remember...
mean(data.hmm_list[[1]]$step, na.rm=T)
sd(data.hmm_list[[1]]$step, na.rm=T)

## can look at the CI for the model
CI(ss_model_list[[1]])
```

```{r}
## then plot the model - this returns a plot of the distribution of step length and and a plot of the turning angle for
## the single state
plot(ss_model_list[[1]])

## you don't get individual plots of movement for the single state model because there is no switching
## between states. if you want a plot of movement for each individual, you can use: plot(data.hmm, compact=FALSE)
```

##Double State Model

```{r}
## give some starting parameter values, because we are trying to identify 2 states now, you need to put in
## the starting parameters that you think are close for each state.

mu0 <- c(mean_step[i], mean_step[i] * 4) # step mean (need one for each state, so should have 2)
sigma0 <- c(sd_step[i], sd_step[i]*3) # step SD 
stepPar0 <- c(mu0,sigma0) ## combine the mean and SD into one object for the model

angleMean0 <- c(mean_angle[i], 0.25) # angle mean
kappa0 <- c(sd_angle[i], 1) # angle concentration
anglePar0 <- c(angleMean0,kappa0)

ds_model_list <- list()

for(i in 1:length(data.hmm_list)){
## run the double state model
doubleStateModel <- fitHMM(data.hmm_list[[i]], nbStates = 2, 
                           stepPar0 = stepPar0,
                           anglePar0 = anglePar0, verbose=2,
                           stepDist = "gamma", angleDist = "vm")

ds_model_list[[i]] <- doubleStateModel

}

##can look at model output
ds_model_list[[5]]

## can look at the CI for the model
CI(ds_model_list[[1]])
```

```{r}
## then plot the model - this returns a plot of the distribution of step length and and a plot of the turning angle for
## the two states determined by the model; it then will give you an output for each indivdiual animal, with the 
## states colored in orange and blue
plot(ds_model_list[[5]])

## this gives you a series of plots for each individual animal. The top plot shows you if they are in state 1 or 2,
## and the bottom two plots gives you the probability that the animal is in state 1 (middle plot) or state 2 (bottom plot)
plotStates(ds_model_list[[5]])
```

##Triple State Model

```{r}
ts_model_list <- list()
## give some starting parameter values, this time we need three means and SDs
for(i in 1:length(data.hmm_list)){
mu0 <- c(50, 200, 500) # step length mean
sigma0 <- c(50, 200, 500) # step length SD
stepPar0 <- c(mu0,sigma0)

angleMean0 <- c(0, pi, 0) # angle mean
kappa0 <- c(0.1, 0.5, 0.1) # angle concentration
anglePar0 <- c(angleMean0,kappa0)

tripleStateModel <- fitHMM(data.hmm_list[[i]], nbStates = 3, # Note, this took about 3 minutes
                          stepPar0 = stepPar0, 
                          anglePar0 = anglePar0, verbose=2,
                                                   stepDist = "gamma", angleDist = "vm")

ts_model_list[[i]] <- tripleStateModel

}

#model output
ts_model_list[[5]]

#model CI
CI(ts_model_list[[1]])

```

```{r}
## plot the model, should have the distribution for all three states now, and will plot the states for each
## individual in three colors
plot(ts_model_list[[1]])

## look at the states and probability that each step is in a specific state
plotStates(ts_model_list[[1]])
```

##Mapview Plots

```{r}
# better plots of the results using mapview
data_list[[1]]$state <- viterbi(ts_model_list[[1]])   # add states as a column in your database
# create an sf object of each step in your database
lns <- Points2Lines(data=data_list[[1]], 
                    date_name="date",
                    id_name="id_yr_seas",
                    byid=FALSE,
                    no_cores=11)  # must change this! Try detectCores()-1
head(lns)

# now for the actual plotting
id2plot <- sample(unique(data_list[[1]]$id_yr_seas), 1)  # grab a random id to plot
mapview(data_list[[1]][data_list[[1]]$id_yr_seas == id2plot,], zcol = 'state',layer.name="points") +   # plot it
  mapview(lns[lns$id_yr_seas == id2plot,],layer.name="lines", zcol='state') # how to get color of lines to match?
```

##Model Selection

```{r}
## AIC model comparison by species loop
#loop being weird- adding in 'i' as model name when run, ran aic for each spp separately below instead
# aic_df <- data.frame(Model = "NA", AIC = "NA", species = "NA")
# 
# for(i in 1:length(ss_model_list)){
# aic_df_sp <- AIC(ss_model_list[[i]], ds_model_list[[i]], ts_model_list[[i]]) %>%
#   mutate(species = spp[[i]])
# 
# aic_df <- rbind(aic_df_sp, aic_df)
# }
# 
# aic_df <- aic_df %>% filter(species != "NA")


aic_df_1 <- AIC(ss_model_list[[1]], ds_model_list[[1]], ts_model_list[[1]]) %>%
  mutate(species = spp[[1]])

aic_df_2 <- AIC(ss_model_list[[2]], ds_model_list[[2]], ts_model_list[[2]]) %>%
  mutate(species = spp[[2]])

aic_df_3 <- AIC(ss_model_list[[3]], ds_model_list[[3]], ts_model_list[[3]]) %>%
  mutate(species = spp[[3]])

aic_df_4 <- AIC(ss_model_list[[4]], ds_model_list[[4]], ts_model_list[[4]]) %>%
  mutate(species = spp[[4]])

aic_df_5 <- AIC(ss_model_list[[5]], ds_model_list[[5]], ts_model_list[[5]]) %>%
  mutate(species = spp[[5]])

aic_df <- rbind(aic_df_1, aic_df_2, aic_df_3, aic_df_4, aic_df_5)

write.table(aic_df, file = "./Code output/AICtable_state models_all spp_greenup.txt",
            sep = ",", quote = FALSE, row.names = FALSE)

```

##Lab Write-Up Figures

```{r}
## to get the figures you need for the first part of your write-up, you can use the plot() and 
## plotStates() functions in the moveHMM package. 

plot(ts_model_list[[1]])

plotStates(ts_model_list[[1]])
```

#Export data for daily activity pattern analysis

```{r}
# add states as a column in species' data list
for(i in 1:length(data_list)){
data_list[[i]]$state <- viterbi(ts_model_list[[i]])   
}

#save data
saveRDS(data_list, "./Code output/allspp_greenup_GPSdata_triplestates.rds")
```



