---
title: "YNP Ungulate dBBMM High Use Areas"
author: "Molly Caldwell"
date: "9/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#load packages
library(sf)
library(mapview)
library(tidyverse)
library(dplyr)
library(purrr)
library(adehabitatHR)
library(move)
library(BBMM)
library(lubridate)
library(raster)
```

#Select bison growing season (spring and summer) areas with range of useage (top 75%, top 50%, lowest 10-25% use) as proxy for grazing areas low-high use
Sum raster values for all individual's dBBMM UD rasters. 

```{r}
#Load dBBMM UDs
files_sp = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*BI.*spring.tif$", full.names = TRUE)

files_su = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*BI.*summer.tif$", full.names = TRUE)

files <- c(files_sp, files_su)

dbbm_stack <- stack(files)

#sum rasters
dbbmm_sum <- calc(dbbm_stack, sum)

#save summed raster
writeRaster(dbbmm_sum, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_SUM_ALLbison_growseas.tif", 
            format="GTiff", overwrite=T)
```

```{r}
#extract areas with top 75% raster values
max_value <- max(dbbmm_sum@data@max)
perc75_min_range <- max_value - (max_value*0.75)

dbbmm_sum_top75perc <- dbbmm_sum >= perc75_min_range

#save extracted raster
writeRaster(dbbmm_sum_top75perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top75perc_ALLbison_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top75perc <- rasterToPolygons(dbbmm_sum_top75perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top75perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top75perc_ALLbison_growseas",
         driver="ESRI Shapefile")
```

```{r}
#extract areas with top 50% raster values
max_value <- max(dbbmm_sum@data@max)
perc50_min_range <- max_value - (max_value*0.50)

dbbmm_sum_top50perc <- dbbmm_sum >= perc50_min_range

#save extracted raster
writeRaster(dbbmm_sum_top50perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top50perc_ALLbison_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top50perc <- rasterToPolygons(dbbmm_sum_top50perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top50perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top50perc_ALLbison_growseas",
         driver="ESRI Shapefile")
```

```{r}
#extract areas with bottom 10-25% raster values
max_value <- max(dbbmm_sum@data@max)
perc10to25_min <- (max_value*0.40)
perc10to25_max <- (max_value*0.50)


dbbmm_sum_low10to25perc <- dbbmm_sum <= perc10to25_max & dbbmm_sum >= perc10to25_min

#save extracted raster
writeRaster(dbbmm_sum_low10to25perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_40to50perc_ALLbison_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_low10to25perc <- rasterToPolygons(dbbmm_sum_low10to25perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_low10to25perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_40to50perc_ALLbison_growseas",
         driver="ESRI Shapefile")
```

#Select bighorn growing season areas with highest usage (top 75% and top 50%)
Sum raster values for all individual's dBBMM UD rasters. 

```{r}
#Load dBBMM UDs
files_sp = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*BH.*spring.tif$", full.names = TRUE)

files_su = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*BH.*summer.tif$", full.names = TRUE)

files <- c(files_sp, files_su)

dbbm_stack <- stack(files)

#sum rasters
dbbmm_sum <- calc(dbbm_stack, sum)

#save summed raster
writeRaster(dbbmm_sum, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_SUM_ALLbighorn_growseas.tif", 
            format="GTiff", overwrite=T)
```

```{r}
#extract areas with top 75% raster values
max_value <- max(dbbmm_sum@data@max)
perc75_min_range <- max_value - (max_value*0.75)

dbbmm_sum_top75perc <- dbbmm_sum >= perc75_min_range

#save extracted raster
writeRaster(dbbmm_sum_top75perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top75perc_ALLbighorn_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top75perc <- rasterToPolygons(dbbmm_sum_top75perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top75perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top75perc_ALLbighorn_growseas",
         driver="ESRI Shapefile")
```

```{r}
#extract areas with top 50% raster values
max_value <- max(dbbmm_sum@data@max)
perc50_min_range <- max_value - (max_value*0.50)

dbbmm_sum_top50perc <- dbbmm_sum >= perc50_min_range

#save extracted raster
writeRaster(dbbmm_sum_top50perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top50perc_ALLbighorn_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top50perc <- rasterToPolygons(dbbmm_sum_top50perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top50perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top50perc_ALLbighorn_growseas",
         driver="ESRI Shapefile")
```

#Select deer growing season areas with highest usage (top 75% and top 50%)
Sum raster values for all individual's dBBMM UD rasters. 

```{r}
#Load dBBMM UDs
files_sp = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*MD.*spring.tif$", full.names = TRUE)

files_su = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*MD.*summer.tif$", full.names = TRUE)

files <- c(files_sp, files_su)

dbbm_stack <- stack(files)

#sum rasters
dbbmm_sum <- calc(dbbm_stack, sum)

#save summed raster
writeRaster(dbbmm_sum, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_SUM_ALLdeer_growseas.tif", 
            format="GTiff", overwrite=T)
```

```{r}
#extract areas with top 75% raster values
max_value <- max(dbbmm_sum@data@max)
perc75_min_range <- max_value - (max_value*0.75)

dbbmm_sum_top75perc <- dbbmm_sum >= perc75_min_range

#save extracted raster
writeRaster(dbbmm_sum_top75perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top75perc_ALLdeer_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top75perc <- rasterToPolygons(dbbmm_sum_top75perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top75perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top75perc_ALLdeer_growseas",
         driver="ESRI Shapefile")
```

```{r}
#extract areas with top 50% raster values
max_value <- max(dbbmm_sum@data@max)
perc50_min_range <- max_value - (max_value*0.50)

dbbmm_sum_top50perc <- dbbmm_sum >= perc50_min_range

#save extracted raster
writeRaster(dbbmm_sum_top50perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top50perc_ALLdeer_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top50perc <- rasterToPolygons(dbbmm_sum_top50perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top50perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top50perc_ALLdeer_growseas",
         driver="ESRI Shapefile")
```

#Select elk growing season areas with highest usage (top 75% and top 50%)
Sum raster values for all individual's dBBMM UD rasters. 

```{r}
#Load dBBMM UDs
files_sp = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*EL.*spring.tif$", full.names = TRUE)

files_su = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*EL.*summer.tif$", full.names = TRUE)

files <- c(files_sp, files_su)

dbbm_stack <- stack(files)

#sum rasters
dbbmm_sum <- calc(dbbm_stack, sum)

#save summed raster
writeRaster(dbbmm_sum, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_SUM_ALLelk_growseas.tif", 
            format="GTiff", overwrite=T)
```

```{r}
#extract areas with top 75% raster values
max_value <- max(dbbmm_sum@data@max)
perc75_min_range <- max_value - (max_value*0.75)

dbbmm_sum_top75perc <- dbbmm_sum >= perc75_min_range

#save extracted raster
writeRaster(dbbmm_sum_top75perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top75perc_ALLelk_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top75perc <- rasterToPolygons(dbbmm_sum_top75perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top75perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top75perc_ALLelk_growseas",
         driver="ESRI Shapefile")
```

```{r}
#extract areas with top 50% raster values
max_value <- max(dbbmm_sum@data@max)
perc50_min_range <- max_value - (max_value*0.50)

dbbmm_sum_top50perc <- dbbmm_sum >= perc50_min_range

#save extracted raster
writeRaster(dbbmm_sum_top50perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top50perc_ALLelk_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top50perc <- rasterToPolygons(dbbmm_sum_top50perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top50perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top50perc_ALLelk_growseas",
         driver="ESRI Shapefile")
```

#Select pronghorn growing season areas with highest usage (top 75% and top 50%)
Sum raster values for all individual's dBBMM UD rasters. 

```{r}
#Load dBBMM UDs
files_sp = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*PR.*spring.tif$", full.names = TRUE)

files_su = list.files("./Code output/dBBMM_UDs/season/", 
                   pattern="*PR.*summer.tif$", full.names = TRUE)

files <- c(files_sp, files_su)

dbbm_stack <- stack(files)

#sum rasters
dbbmm_sum <- calc(dbbm_stack, sum)

#save summed raster
writeRaster(dbbmm_sum, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_SUM_ALLpronghorn_growseas.tif", 
            format="GTiff", overwrite=T)
```

```{r}
#extract areas with top 75% raster values
max_value <- max(dbbmm_sum@data@max)
perc75_min_range <- max_value - (max_value*0.75)

dbbmm_sum_top75perc <- dbbmm_sum >= perc75_min_range

#save extracted raster
writeRaster(dbbmm_sum_top75perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top75perc_ALLpronghorn_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top75perc <- rasterToPolygons(dbbmm_sum_top75perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top75perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top75perc_ALLpronghorn_growseas",
         driver="ESRI Shapefile")
```

```{r}
#extract areas with top 50% raster values
max_value <- max(dbbmm_sum@data@max)
perc50_min_range <- max_value - (max_value*0.50)

dbbmm_sum_top50perc <- dbbmm_sum >= perc50_min_range

#save extracted raster
writeRaster(dbbmm_sum_top50perc, filename = "./Code output/dBBMM_UDs/UD_SUMS/dBBMM_top50perc_ALLpronghorn_growseas.tif", 
            format="GTiff", overwrite=T)

#convert to polygons and save as shapefile
poly_top50perc <- rasterToPolygons(dbbmm_sum_top50perc, fun = function(x){x > 0},
                                   dissolve = TRUE)

writeOGR(obj = poly_top50perc, dsn = "~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dBBMM_UDs/UD_SUMS", 
         layer = "dBBMM_UD_top50perc_ALLpronghorn_growseas",
         driver="ESRI Shapefile")
```