---
title: "YNP ungulates proposal talk examples"
author: "Molly Caldwell"
date: "10/22/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#load packages
library(tidyverse)
library(ggridges)
library(Ostats)
library(parallel)
library(raster)
library(sf)
library(mapview)
library(moveVis)
library(lubridate)

#set wd
setwd("~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```


```{r}
#load all gps data
data <- readRDS("./Data/GPS data/Cleaned/allspp_cleanedGPSdata_seasonal_3.2022.rds")

```

#Bison movement example 1
2 individual bison that migrate during spring that exhibit different paths.

```{r}
#filter data to bison
bison <- data %>% 
  filter(id_yr_seas %in% c("BI131_2021_spring", "BI138_2021_spring"))

#filter data to early spring 
bison_es <- bison %>% 
  filter(date < as.POSIXct("2021-06-01 00:00:00", format = "%Y-%m-%d %H:%M:%S"))

#project data
bison_es <- st_transform(bison_es, crs = CRS("+proj=longlat +ellps=WGS84"))

#separate geometry to x y columns
bison_es <- bison_es %>% 
   mutate(x = unlist(purrr::map(bison_es$geometry,1)),
           y = unlist(purrr::map(bison_es$geometry,2)))

bison_es$geometry <- NULL

#order by id and timestamp
bison_es <- bison_es %>% 
  group_by(id_yr_seas) %>% 
  arrange(date, .by_group = T) %>% 
  dplyr::select(c("id_yr_seas", "x", "y", "date")) %>% 
  ungroup()

#rename ids
bison_es <- bison_es %>% 
  mutate(id_yr_seas = if_else(id_yr_seas == "BI131_2021_spring", "Bison 131",
                              "Bison 138"))

#transform data to move class object
bison_m <- df2move(bison_es, proj = CRS("+proj=longlat +ellps=WGS84"), x = "x", y = "y",
                   time = "date", track_id = "id_yr_seas")



#align data to a uniform time scale
u_bison <- align_move(bison_m, res = 2, unit = "hours")

# create spatial frames with a OpenStreetMap watercolour map
frames <- frames_spatial(u_bison, path_colours = c("goldenrod1", "goldenrod4"),
                         map_service = "osm", map_type = "topographic", alpha = 0.5, 
                         tail_colour = c("grey"), tail_size = 0.8,
                         trace_show = T, trace_colour = c("grey"),
                         equidistant = F) %>% 
  add_labels(x = "Longitude", y = "Latitude") %>% 
  add_northarrow() %>% 
  add_scalebar() %>% 
  add_timestamps(type = "label") %>% 
  add_progress()


frames[[5]] # preview one of the frames, e.g. the 100th frame
```

#Bison movement example 2
2 individual bison that show collective movement.

```{r}
#filter data to contact days
bison_c <- bison_es %>% 
  filter(date > as.POSIXct("2021-05-13 00:00:00", format = "%Y-%m-%d %H:%M:%S") &
           date < as.POSIXct("2021-05-16 00:00:00", format = "%Y-%m-%d %H:%M:%S"))


#transform data to move class object
bison_m2 <- df2move(bison_c, proj = CRS("+proj=longlat +ellps=WGS84"), x = "x", y = "y",
                   time = "date", track_id = "id_yr_seas")



#align data to a uniform time scale
u_bison2 <- align_move(bison_m2, res = 30, unit = "mins")

# create spatial frames with a OpenStreetMap watercolour map
frames2 <- frames_spatial(u_bison2, path_colours = c("goldenrod1", "goldenrod4"),
                         map_service = "osm", map_type = "topographic", alpha = 0.5, 
                         tail_colour = c("grey"), tail_size = 0.8,
                         trace_show = T, trace_colour = c("grey"),
                         equidistant = F) %>% 
  add_labels(x = "Longitude", y = "Latitude") %>% 
  add_northarrow() %>% 
  add_scalebar() %>% 
  add_timestamps(type = "label") %>% 
  add_progress()

#export gif
animate_frames(frames2, out_file = "bison_example2.gif")

```

#Bison habitat use
ridgeline plot showing individual variation habitat use

```{r}
#load gps points with habitat data
hab_data <- readRDS("./Code output/individual_habitat_metrics/allindiv_gpspts_habextract.rds")

#filter by bison and spring 2021
bison_hab <- hab_data %>% 
  filter(species == "bison" & season == "spring" & yr == 2021) %>% 
  mutate(cid = factor(cid, levels = c("BI047", "BI070", "BI100", "BI127", "BI132", "BI134", "BI137",  "BI139", "BI140", "BI138", "BI131")))

```

```{r}
#ridgeline plots
library(ggridges)
library(ggpubr)
library(RColorBrewer)

col <- heat.colors(11)

#graph forbgrass biomass by indiv
ggplot(data = bison_hab, aes(x = forbgrass_biomass, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 5, alpha = 0.2)) +
  scale_fill_manual(values = c("#999999", "#CC79A7", "#56B4E9",
        "firebrick4", "#009E73", "#F0E442", "magenta4", "#D55E00", 
        "#0072B2","chocolate4", "goldenrod")) +
  xlab("Forb and grass biomass (lbs/acre)") +
  ylab("")+
  theme(legend.position = "none")

#graph elevation by indiv
ggplot(data = bison_hab, aes(x = elev, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 5, alpha = 0.2)) +
  scale_fill_manual(values = c("#999999", "#CC79A7", "#56B4E9",
        "firebrick4", "#009E73", "#F0E442", "magenta4", "#D55E00", 
        "#0072B2","chocolate4", "goldenrod")) +
  xlab("Forb and grass biomass (lbs/acre)") +
  xlab("Elevation (m)") +
  ylab("")+
  theme(legend.position = "none")

```


#Bison and pronghorn movement animation example
2 individual bison and pronghorn that migrate during spring that exhibit different paths.

```{r}
#filter data to example individuals
bi_ph <- data %>% 
  filter(id_yr_seas %in% c("BI131_2021_spring", "BI138_2021_spring",
                           "PR27_2021_spring", "PR24_2021_spring"))

#filter data to early spring 
bi_ph_es <- bi_ph %>% 
  filter(date < as.POSIXct("2021-06-01 00:00:00", format = "%Y-%m-%d %H:%M:%S"))

#project data
bi_ph_es <- st_transform(bi_ph_es, crs = CRS("+proj=longlat +ellps=WGS84"))

#separate geometry to x y columns
bi_ph_es <- bi_ph_es %>% 
   mutate(x = unlist(purrr::map(bi_ph_es$geometry,1)),
           y = unlist(purrr::map(bi_ph_es$geometry,2)))

bi_ph_es$geometry <- NULL

#rename ids
bi_ph_es <- bi_ph_es %>% 
  mutate(id_yr_seas = if_else(id_yr_seas == "BI131_2021_spring", "Bison 131",
                          if_else(id_yr_seas == "BI138_2021_spring","Bison 138",
                          if_else(id_yr_seas == "PR27_2021_spring", "Pronghorn 27",
                                  "Pronghorn 24"))))

#order by id and timestamp
bi_ph_es <- bi_ph_es %>% 
  group_by(id_yr_seas) %>% 
  arrange(date, .by_group = T) %>% 
  dplyr::select(c("id_yr_seas", "x", "y", "date")) %>% 
  ungroup()



#transform data to move class object
bph_m <- df2move(bi_ph_es, proj = CRS("+proj=longlat +ellps=WGS84"), x = "x", y = "y",
                   time = "date", track_id = "id_yr_seas")



#align data to a uniform time scale
u_bph <- align_move(bph_m, res = 2, unit = "hours")

# create spatial frames with a OpenStreetMap watercolour map
frames3 <- frames_spatial(u_bph, path_colours = c("goldenrod1", "goldenrod4",
                                                 "hotpink1", "indianred"),
                         map_service = "osm", map_type = "topographic", alpha = 0.5, 
                         tail_colour = c("grey"), tail_size = 0.8,
                         trace_show = T, trace_colour = c("grey"),
                         equidistant = F) %>% 
  add_labels(x = "Longitude", y = "Latitude") %>% 
  add_northarrow() %>% 
  add_scalebar() %>% 
  add_timestamps(type = "label") %>% 
  add_progress()


#export gif
animate_frames(frames3, out_file = "bison_pronghorn_example3.gif")
```

#Bison and pronghorn habitat graphs
ridgeline plot showing individual variation habitat use

```{r}
#load gps points with habitat data
hab_data <- readRDS("./Code output/individual_habitat_metrics/allindiv_gpspts_habextract.rds")

#filter by example bison and pronghorn individuals
bph_hab <- hab_data %>% 
  filter(id_yr_seas %in% c("BI131_2021_spring", "BI138_2021_spring",
                           "PR27_2021_spring", "PR24_2021_spring")) %>% 
  mutate(cid = factor(cid, levels = c("PR27", "PR24", "BI138", "BI131")))

```

```{r}
#ridgeline plots
library(ggridges)
library(ggpubr)
library(RColorBrewer)

col <- heat.colors(11)

#graph forbgrass biomass by indiv
ggplot(data = bph_hab, aes(x = forbgrass_biomass, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("indianred", "hotpink1", "goldenrod4", "goldenrod1")) +
  xlab("Forb and grass biomass (lbs/acre)") +
  ylab("")+
  theme(legend.position = "none")


#graph elevation by indiv
ggplot(data = bph_hab, aes(x = elev, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("indianred", "hotpink1", "goldenrod4", "goldenrod1")) +
  xlab("Elevation (m)") +
  ylab("")+
  theme(legend.position = "none")

```

```{r}
#BI 138 and PR24 graph
hab_sub1 <- bph_hab %>% 
  filter(id_yr_seas %in% c("BI138_2021_spring", "PR27_2021_spring"))

#graph forbgrass biomass by indiv
ggplot(data = hab_sub1, aes(x = forbgrass_biomass, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("indianred", "goldenrod4")) +
  xlab("Forb and grass biomass (lbs/acre)") +
  ylab("")+
  theme(legend.position = "none")


#graph elevation by indiv
ggplot(data = hab_sub1, aes(x = elev, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("indianred", "goldenrod4")) +
  xlab("Elevation (m)") +
  ylab("")+
  theme(legend.position = "none")
```

```{r}
#BI 138, PR27 and PR24 graph
hab_sub2 <- bph_hab %>% 
  filter(id_yr_seas %in% c("BI138_2021_spring", "PR27_2021_spring", "PR24_2021_spring")) %>% 
  mutate(cid = ordered(cid, levels = c("PR24", "PR27", "BI138")))

#graph forbgrass biomass by indiv
ggplot(data = hab_sub2, aes(x = forbgrass_biomass, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("hotpink1", "indianred", "goldenrod4")) +
  xlab("Forb and grass biomass (lbs/acre)") +
  ylab("")+
  theme(legend.position = "none")


#graph elevation by indiv
ggplot(data = hab_sub2, aes(x = elev, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("hotpink1", "indianred", "goldenrod4")) +
  xlab("Elevation (m)") +
  ylab("")+
  theme(legend.position = "none")
```



#Yellowstone group interactions graph

```{r}
#generate data on number of groups and number species/group
group_df <- data.frame(Nspecies = c(2,3,4),
                        Ngroups = c(13932, 343, 8))


ggplot(data = group_df, aes(x = Nspecies, y = Ngroups, col = Nspecies, size = Nspecies)) +
  geom_point() +
  theme_bw()

# Compute percentages
group_df$fraction = group_df$Ngroups / sum(group_df$Ngroups)

# Compute the cumulative percentages (top of each rectangle)
group_df$ymax = cumsum(group_df$fraction)

# Compute the bottom of each rectangle
group_df$ymin = c(0, head(group_df$ymax, n=-1))

# Compute label position
group_df$labelPosition <- (group_df$ymax + group_df$ymin) / 2

#label
group_df$label <- paste0(group_df$Nspecies, "\n value: ", group_df$Ngroups)
 
# Make the plot
ggplot(group_df, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Nspecies)) +
     geom_rect() +
  #scale_fill_brewer(palette=4) +
     coord_polar(theta="y") + # Try to remove that to understand how the chart is built initially
     xlim(c(2, 4)) + # Try to remove that to see how to make a pie chart
  theme_void() +
  theme(legend.position = "none")

```


#HR overlap greenup

Finding individuals of different species with high amount of HR overlap

```{r}
#load proportion overlap data
prop_ol <- read.csv("~/UWyo/PhD project/YNP-ungulate-niche-proj/Code output/dbbmm_HR_overlap_greenup_allindiv.csv")

#add years of HRs
prop_ol <- prop_ol %>% 
  rowwise() %>% 
  mutate(year1 = str_split(id1, "_")[[1]][3]) %>% 
  mutate(year2 = str_split(id2, "_")[[1]][3])

#filter by species different and years same
ol_spp <- prop_ol %>% 
  filter((year1 == year2) & (spp1 != spp2)) %>% 
  arrange(prop_overlap)
```

```{r}
#filter by example bison and pronghorn individuals
bph_hab <- hab_data %>% 
  filter(id_yr_seas %in% c("BI118_2018_spring", "PR09_2018_spring", "BI120_2018_spring")) %>% 
  mutate(cid = ordered(cid, levels = c("PR09", "BI118", "BI120")))
```

```{r}
#graph forbgrass biomass by indiv
ggplot(data = bph_hab, aes(x = forbgrass_biomass, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("#0072B2", "#009E73", "#D55E00")) +
  xlab("Forb and grass biomass (lbs/acre)") +
  ylab("")+
  theme(legend.position = "none")


#graph elevation by indiv
ggplot(data = bph_hab, aes(x = elev, y = cid, fill = cid)) +
  geom_density_ridges(aes(scale = 4, alpha = 0.2)) +
  scale_fill_manual(values = c("#0072B2", "#009E73", "#D55E00")) +
  xlab("Elevation (m)") +
  ylab("")+
  theme(legend.position = "none")
```





