---
title: "YNP_ungulate_indiv_bufferHRs"
author: "Molly Caldwell"
date: '2022-06-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
knitr::opts_knit$set(root.dir = "~/UWyo/PhD project/YNP-ungulate-niche-proj/")
```

```{r}
#packages
library(sf)
library(raster)
library(mapview)
library(tidyverse)
library(adehabitatHR)
library(move)
library(lubridate)
library(parallel)
library(tidyverse)
library(multidplyr)
```

#Load GPS Data

```{r}
#load seasonal GPS data with 2 hour fixes for all species
data <- readRDS("./Data/GPS data/Cleaned/allspp_cleanedGPSdata_seasonal_3.2022.rds")

data <- st_as_sf(data)

#add date only column
data <- data %>%
  mutate(date_only = format(date, format = "%Y-%m-%d"))

#remove extra date column
data$date2 <- NULL
```

#Buffer home range- daily
Turns gps point path into lines and puts 200-m buffer around them.
Buffer HRs by id_yr_seas and day.

```{r}
#create list to save line data
lns_list <- list()

##setup cluster to run on multiple cores
  #set number of cores to use (1 less than you have)
no_cores <- detectCores()-1
# Setup cluster
clust <- makeCluster(no_cores) 
# export the objects you need for your calculations from your environment to each node's environment
clusterExport(clust, varlist=c("data"))

u_id <- unique(data$id_yr_seas)

for(i in 1:length(u_id)){
  print(paste0("i: ", i))
  
  data_id <- data %>% filter(id_yr_seas == u_id[i])
  
  dt <- unique(data_id$date_only)
  
  lns <- do.call(c, lapply(1:length(dt), function(e){

  return(st_cast(st_combine(data_id[data_id$date_only == dt[e],]), "LINESTRING"))

}))
  
  lns <- data.frame(id_yr_seas = u_id[i],
                  firstdate=do.call(c, lapply(dt, 
                function(e){min(data_id$date[data_id$date_only == e], na.rm=TRUE)})),
              lastdate=do.call(c, lapply(dt, 
              function(e){max(data_id$date[data_id$date_only == e], na.rm=TRUE)})),
              geometry=lns)
  
  
  lns_list[[i]] <- lns
  
}
 
lns <- data.frame(Reduce(rbind, lns_list))

lns <- st_as_sf(lns, sf_column_name = "geometry")

#remove observation with only one pt/day
lns <- lns %>%
  group_by(id_yr_seas, firstdate) %>%
  mutate(length_geom = length(geometry[[1]])) %>%
  filter(length_geom > 2)

# now calculate the Home range polygon for each line string
HRs_day <- st_buffer(lns, dist = 200)    #200m buffer

stopCluster(clust)   #stop the parallelization framework

#add species to HRs
HRs_day <- HRs_day %>%
  mutate(species = if_else(grepl("BI", id_yr_seas), "bison",
                if_else(grepl("BH", id_yr_seas), "bighorn",
                if_else(grepl("MD", id_yr_seas), "deer",
                if_else(grepl("EL", id_yr_seas), "elk",
                if_else(grepl("PR", id_yr_seas), "pronghorn", "NA"))))))

#individual HR area summary- takes awhile
cluster <- new_cluster(no_cores)
cluster_library(cluster, packages = c("dplyr", "sf"))
cluster_copy(cluster, "HRs_day")

HRs_day <- HRs_day %>% partition(cluster) %>%
  mutate(HR_area = st_area(geometry)) %>%
  mutate(HR_area = as.numeric(HR_area)/1000000) %>% #convert to sq km
  collect()

#remove length geom column
HRs_day$length_geom <- NULL

#save as rds object
saveRDS(HRs_day, "./Code output/individual_buffer_HRs/indiv_buffHR_day.rds")
```

#Buffer home range- weekly (5 days)
Turns gps point path into lines and puts 200-m buffer around them.
Buffer HRs by id_yr_seas and week.

```{r}
#create 5-day grouping variable per id_yr_seas
data <- data %>% group_by(id_yr_seas) %>%
  mutate(week = cut(as.Date(date), breaks = "5 days", include.lowest = T))
  
```

```{r}
#create list to save line data
lns_list <- list()

##setup cluster to run on multiple cores
  #set number of cores to use (1 less than you have)
no_cores <- detectCores()-1
# Setup cluster
clust <- makeCluster(no_cores) 
# export the objects you need for your calculations from your environment to each node's environment
clusterExport(clust, varlist=c("data"))

u_id <- unique(data$id_yr_seas)

for(i in 1:length(u_id)){
  print(paste0("i: ", i))
  
  data_id <- data %>% filter(id_yr_seas == u_id[i])
  
  dt <- unique(data_id$week)
  
  lns <- do.call(c, lapply(1:length(dt), function(e){

  return(st_cast(st_combine(data_id[data_id$week == dt[e],]), "LINESTRING"))

}))
  
  lns <- data.frame(id_yr_seas = u_id[i],
                  firstdate=do.call(c, lapply(dt, 
                function(e){min(data_id$date[data_id$week == e], na.rm=TRUE)})),
              lastdate=do.call(c, lapply(dt, 
              function(e){max(data_id$date[data_id$week == e], na.rm=TRUE)})),
              geometry=lns)
  
  
  lns_list[[i]] <- lns
  
}
 
lns <- data.frame(Reduce(rbind, lns_list))

lns <- st_as_sf(lns, sf_column_name = "geometry")

#remove observation with only one pt/day
lns <- lns %>%
  group_by(id_yr_seas, firstdate) %>%
  mutate(length_geom = length(geometry[[1]])) %>%
  filter(length_geom > 2)

# now calculate the Home range polygon for each line string
HRs_week <- st_buffer(lns, dist = 200)    #200m buffer

#add species to HRs
HRs_week <- HRs_week %>%
  mutate(species = if_else(grepl("BI", id_yr_seas), "bison",
                if_else(grepl("BH", id_yr_seas), "bighorn",
                if_else(grepl("MD", id_yr_seas), "deer",
                if_else(grepl("EL", id_yr_seas), "elk",
                if_else(grepl("PR", id_yr_seas), "pronghorn", "NA"))))))

#individual HR area summary
HRs_week <- HRs_week %>%
  mutate(HR_area = st_area(geometry)) %>%
  mutate(HR_area = as.numeric(HR_area)/1000000) #convert to sq km

#remove length geom column
HRs_week$length_geom <- NULL

stopCluster(clust)   #stop the parallelization framework

#save as rds object
saveRDS(HRs_week, "./Code output/individual_buffer_HRs/indiv_buffHR_week.rds")
```

#Buffer home range- monthly 
Turns gps point path into lines and puts 200-m buffer around them.
Buffer HRs by id_yr_seas and month.

```{r}
#create month grouping variable per id_yr_seas
data <- data %>% group_by(id_yr_seas) %>%
  mutate(month = cut(as.Date(date), breaks = "1 month", include.lowest = T))
  
```

```{r}
#create list to save line data
lns_list <- list()

##setup cluster to run on multiple cores
  #set number of cores to use (1 less than you have)
no_cores <- detectCores()-1
# Setup cluster
clust <- makeCluster(no_cores) 
# export the objects you need for your calculations from your environment to each node's environment
clusterExport(clust, varlist=c("data"))

u_id <- unique(data$id_yr_seas)

for(i in 1:length(u_id)){
  print(paste0("i: ", i))
  
  data_id <- data %>% filter(id_yr_seas == u_id[i])
  
  dt <- unique(data_id$month)
  
  lns <- do.call(c, lapply(1:length(dt), function(e){

  return(st_cast(st_combine(data_id[data_id$month == dt[e],]), "LINESTRING"))

}))
  
  lns <- data.frame(id_yr_seas = u_id[i],
                  firstdate=do.call(c, lapply(dt, 
                function(e){min(data_id$date[data_id$month == e], na.rm=TRUE)})),
              lastdate=do.call(c, lapply(dt, 
              function(e){max(data_id$date[data_id$month == e], na.rm=TRUE)})),
              geometry=lns)
  
  
  lns_list[[i]] <- lns
  
}
 
lns <- data.frame(Reduce(rbind, lns_list))

lns <- st_as_sf(lns, sf_column_name = "geometry")

#remove observation with only one pt/day
lns <- lns %>%
  group_by(id_yr_seas, firstdate) %>%
  mutate(length_geom = length(geometry[[1]])) %>%
  filter(length_geom > 2)

# now calculate the Home range polygon for each line string
HRs_month <- st_buffer(lns, dist = 200)    #200m buffer

#add species to HRs
HRs_month <- HRs_month %>%
  mutate(species = if_else(grepl("BI", id_yr_seas), "bison",
                if_else(grepl("BH", id_yr_seas), "bighorn",
                if_else(grepl("MD", id_yr_seas), "deer",
                if_else(grepl("EL", id_yr_seas), "elk",
                if_else(grepl("PR", id_yr_seas), "pronghorn", "NA"))))))

#individual HR area summary
HRs_month <- HRs_month %>%
  mutate(HR_area = st_area(geometry)) %>%
  mutate(HR_area = as.numeric(HR_area)/1000000) #convert to sq km

#remove length geom column
HRs_month$length_geom <- NULL

stopCluster(clust)   #stop the parallelization framework

#save as rds object
saveRDS(HRs_month, "./Code output/individual_buffer_HRs/indiv_buffHR_month.rds")
```

#Buffer home range- seasonly 
Turns gps point path into lines and puts 200-m buffer around them.
Buffer HRs by id_yr_seas.

```{r}
data <- data[1:10000,]
##setup cluster to run on multiple cores
  #set number of cores to use (1 less than you have)
no_cores <- detectCores()-1
# Setup cluster
clust <- makeCluster(no_cores) 
# export the objects you need for your calculations from your environment to each node's environment
clusterExport(clust, varlist=c("data"))

u <- unique(data$id_yr_seas)
 
lns <- do.call(c, lapply(1:length(u), function(e){

  return(st_cast(st_combine(data[data$id_yr_seas == u[e],]), "LINESTRING"))

}))

lns <- data.frame(id_yr_seas = u,
                  firstdate=do.call(c, lapply(u, 
                function(e){min(data$date[data$id_yr_seas == e], na.rm=TRUE)})),
              lastdate=do.call(c, lapply(u, 
              function(e){max(data$date[data$id_yr_seas == e], na.rm=TRUE)})),
              geometry=lns)

lns <- st_as_sf(lns, sf_column_name = "geometry")

# now calculate the Home range polygon for each line string
HRs_season <- st_buffer(lns, dist = 200)    #200m buffer

#add species to HRs
HRs_season <- HRs_season %>%
  mutate(species = if_else(grepl("BI", id_yr_seas), "bison",
                if_else(grepl("BH", id_yr_seas), "bighorn",
                if_else(grepl("MD", id_yr_seas), "deer",
                if_else(grepl("EL", id_yr_seas), "elk",
                if_else(grepl("PR", id_yr_seas), "pronghorn", "NA"))))))

#individual HR area summary
HRs_season <- HRs_season %>%
  mutate(HR_area = st_area(geometry)) %>%
  mutate(HR_area = as.numeric(HR_area)/1000000) #convert to sq km


stopCluster(clust)   #stop the parallelization framework

#save as rds object
saveRDS(HRs_season, "./Code output/individual_buffer_HRs/indiv_buffHR_season.rds")
```

#Buffer home range- annually 
Merge seasonal ranges per id and year for individuals with entire yr of data.

```{r}
#load seasonal HRs
HRs_season <- readRDS("./Code output/individual_buffer_HRs/indiv_buffHR_season.rds")

#split id_yr_seas into 3 variables
HRs_season <- HRs_season %>% group_by(id_yr_seas) %>%
  mutate(cid = strsplit(id_yr_seas, "_")[[1]][1]) %>%
   mutate(yr = strsplit(id_yr_seas, "_")[[1]][2]) %>%
   mutate(season = strsplit(id_yr_seas, "_")[[1]][3]) %>%
  ungroup()

#filter by only cids and years with all seasons- 192 id yrs have full yr data
HRs_season_f <- HRs_season %>%
  group_by(cid, yr) %>%
  filter(n() == 4)
```

```{r}
#merge seasonal HR polygons by cid and year
HRs_annual <- HRs_season_f %>%
  group_by(cid, yr) %>%
  summarize(geometry = st_union(geometry), .groups = "keep")

#individual HR area summary
HRs_annual <- HRs_annual %>%
  mutate(HR_area = st_area(geometry)) %>%
  mutate(HR_area = as.numeric(HR_area)/1000000) #convert to sq km

#save as rds object
saveRDS(HRs_annual, "./Code output/individual_buffer_HRs/indiv_buffHR_annual.rds")
```

